### 2022-09-05 products/paper/main.qmd

---
title: Grade Retention and Mixed Race
author: 
  - Aaron Gullickson, *University of Oregon*
abstract: This is a test abstract
thanks: The corresponding author can be reached at aarong@uoregon.edu. All code and data for this project are available at ...
format:
  pdf:
    documentclass: scrartcl
    papersize: letter
    fig-height: 4
    fig-pos: t
    keep-tex: false
  docx:
    toc: false
bibliography: ../project.bib
execute: 
  eval: true
  echo: false
  warning: false
  error: false
editor: source
---

```{r setup}
#| include: false
library(here)
source(here("analysis","check_packages.R"))
source(here("analysis","useful_functions.R"))
```

```{r load-data}
load(here("analysis","output","acs.RData"))
load(here("analysis","output","coef_table.RData"))
load(here("analysis","output","monoracial_results.RData"))
load(here("analysis","output","tbl_resources.RData"))
# change names of models for display
coef_table <- coef_table %>%
  mutate(model=factor(model,
                      levels=c("crude","fixed effects","cultural resources",
                               "parent resources"),
                      labels=c("Raw","Baseline model","Cultural model",
                               "Full model")))
# create design object
acs_design <- acs %>%
  as_survey_design(ids = cluster, weights = perwt)
```

# Introduction

The experiences of mixed race individuals will play an important role in the future of race in the US. A rise in interracial marriage and coupling since the 1960s has resulted in a "biracial baby boom" of individuals whose parents cross racial lines [@root_racially_1992]. This demographic phenomenon may have a potentially transformative effect on racial boundaries [@hochschild_creating_2012; @alba_great_2020]. Although race mixing has always been a feature of US history [@morning_who_2000; @gullickson_choosing_2011], this biracial boom is notable for its demographic size and for occurring in a post-Civil Rights era in which historical norms governing how to handle racial mixture may be shifting. Understanding the transformative potential of this biracial baby boom is contingent on developing a better understanding of how these biracial individuals identify their own race and how they are sorted into a prevailing racial hierarchy that reinforces a system of racial inequality.

Extensive scholarship documents the identification and classification decisions of multiracial individuals [@xie_racial_1997; @roth_end_2005a; @brunsma_interracial_2005; @herman_forced_2004; @holloway_place_2009] <!-- TODO MORE HERE -->. This research has generally identified a surprising degree of fluidity and choice in classification decisions, in the context of historical norms of racial classification <!-- TODO CHECK -->. On the other hand, only a smattering of research has documented how the life chances of biracial individuals compare to their monoracial peers [@kao_racial_1999; @campbell_multiracial_2009; @herman_blackwhiteother_2009]. The lack of work on this topic, despite its acknowledged relevance, results from two important limitations in most available data sources.

First, using direct identification of biracial respondents on data sources such as the Census or large scale surveys is problematic. Research on the identification and classification of biracial individuals indicates that individuals whose parents belong to different races will not always or consistently identify with both races [@doyle_are_2007; @liebler_america_2017; @gullickson_racial_2019]. Furthermore, social status and social mobility has been shown to predict racial identification generally [@saperstein_racial_2012; @saperstein_mulatto_2013] and social status has been shown to be related to the identification choices of biracial individuals in particular [@roth_end_2005a; @davenport_role_2016]. Thus, differences in outcome may in some part produce differences in identification rather than the other way around when examining self-identified biracial individuals.

This problem of reverse causation can be largely addressed when researchers have access to the racial identification of a respondent's biological parents. While such information is available on some large scale longitudinal surveys, researchers then run into a second problem of sample size. The samples of biracial respondents in such cases are generally far too small to make reliable statistical inferences. This limitation is particularly severe for biracial individuals with two non-white parents.

When children reside with both biological parents, biracial individuals can be identified in large scale data sources such as the Census. Researchers have long utilized the ability to identify such biracial children to examine the classification choices made on behalf of these children [@xie_racial_1997; @roth_end_2005a]. Because of their age, it is more difficult to analyze the outcomes of such children due to the simple fact that they haven't had many measurable life outcomes yet.

However, one important outcome, grade retention, is measurable and important even for young children. Being held back in school is a relatively common occurrence that has consequences for later educational transitions and is disproportionately applied by race <!-- TODO CITATIONS -->. Grade retention is arguably one of the earliest stratifiers of life chances that an individual will experience and within it we can clearly see the imprint of the racial hierarchy. In this article, I use data from the American Community Survey to examine the likelihood of being held back in school for biracial children in comparison to their monoracial peers. Pooling data from 2010 to 2019, I am able to draw a large enough sample to make reliable statistical inferences for most biracial groups, including those with two non-white parents.

The results provide a window into the placement of biracial children within an existing system of racial inequality in our educational systems. <!-- TODO  PITHY SUMMARY -->

# Where do Biracials Fit?

How might we expect the outcomes for biracial individuals do differ from their monoracial peers? In every case, the relevant comparison is how a biracial person fares relative to the two constituent groups with which their parents identify. For example, in the case of an individual with one Black and one White parent, we would compare the average outcomes for such individuals to the average outcomes for individuals with two White and two Black parents. The outcome for the Black/White group may be similar to one of the monoracial groups, it may be somewhere in between the outcomes for the two monoracial groups, or it may be higher or lower than either of the monoracial groups.

To better understand where biracial individuals might fit, it is necessary to understand what is producing a contemporary system of racial inequality overall. Here, I distinguish three mechanisms that may help us understand the biracial position within this racial hierarchy. First, I consider arguments for the racial hierarchy that focus on the effects of racial discrimination at both an individual and systemic level. Second, I consider arguments regarding different distributions of resources across racial groups that produce different average outcomes. Third, I consider arguments regarding how the nature of hybridity itself may affect outcomes.

## Discrimination and Classification

Discrimination and bias by race play an important role in overall levels of racial inequality. Discrimination need not be consciously recognized by actors, but may instead reflect cognitive biases and schema that individuals unknowningly rely upon and reinforce in social interactions <!-- TODO citations, may be Pager, Brubaker? -->. Such discrimination will be most consequential when performed by important gatekeepers of opportunity such as teachers, health care workers, and employers. Discrimination need not just occur on an individual basis, but can also be systemic in nature when biases are built into organizations and institutional systems <!-- TODO citations, Tilly? Bonilla-Silva, Pager again? -->. In our current "colorblind" regime, such systems may be formally race neutral but produce disparate impacts by race <!-- TODO CITE Bonilla-Silva? -->.

The experiences of biracial individuals within such a system of discrimination depend almost entirely on how they are classified by others. To the extent that biracial individuals are classified as more like one of their monoracial constituent groups than the other, they will receive the same treatment and reap similar rewards or suffer similar penalties.

The strongest historical precedent for understanding such classification in the US is the "one drop rule" norm concerning Black ancestry. According to the one-drop rule, individuals of known Black ancestry are considered exclusively Black, regardless of other ancestries[@davis_who_1991]. @davis_who_1991 generalizes this practice into the concept of *hypodescent* in which individuals with mixed racial ancestry will be classified as members of the lower status group. By comparison a pattern of *hyperdescent* implies that individuals of mixed racial ancestry will be classified as members of the higher status group. This pattern of hyperdescent has historically governed the experiences of individuals with American Indian ancestry <!-- TODO CITE -->.

Although the hypodescent/hyperdescent paradigm is often used to theoretically frame the possible contemporary experiences of multiracial individuals <!-- TODO citations -->, there are reasons to be cautious about the generalizability of this paradigm to both the contemporary and historical situation. Hypodescent and hyperdescent are observed in two specific historical cases involving Black ancestry and American Indian ancestry, respectively. As @iverson_regimes_2022 note, these two cases could just as parsimoniously be explained by a "dominance" model which ranks ancestries by their tendencies to be supercessive or recessive in determining identification and classification. In this framework, within the historical "racial triangle" of American colonialism, Black ancestry is supercessive, Indigenous ancestry is recessive, and White ancestry falls somewhere in between.

The development of these particular norms may also reflect the historical regime in which they were developed [@gullickson_choosing_2011; @iverson_regimes_2022]. In this case, neither the hypodescent/hyperdescent or dominance paradigm may help us to fully understand the experiences of those of Latino and Asian ancestry due to more recent processes of racialization. Second, the historical norms for all groups may also be breaking up or transforming in the post-Civil Rights era <!-- TODO CITE -->.

If outcomes for biracial individuals all closer to one of their constituent groups, then we may have stronger evidence about such norms of classification. However, what if those outcomes fall in-between the two groups? How might classification and discrimination produce such an in-between status? There are several possibilities here.

First, racial ambiguity in appearance may lead to more inconsistent classification into single race categories across observers. Even if norms consistently indicate a certain type of classification, correct identification by observers is still a probabilistic phenomenon in all social interactions and likely to be more heterogeneous for biracial individuals. As a result, biracial individuals will experience a lower sum total of discrimination and bias affecting the lower status group in the case of hypodescent and will not reap the same rewards as a member of the higher status group in the case of hyperdescent.

Second, even when "correctly" classified, greater physical ambiguity may allow biracial individuals to avoid the worst instances of bias and discrimination because they do not conform to the ideal type for the group facing the most discrimination. In other words, biracial people's experience of discrimination may be "softened" because their ambiguous physical appearance provokes less racial antipathy. The US in fact already has a long history of this kind of "softening." Because of the one-drop rule, Black individuals vary substantially in skin tone and those individuals with a lighter skin tone have better outcomes [@keith_skin_1991; @hughes_significance_1990; @monk_skin_2014] and are viewed more favorably by both Blacks and Whites [@maddox_cognitive_2002a]. Thus, light skin moderates the stigma of Blackness even among monoracial individuals.

Third, it is also possible that biracial individuals may be classified into an ambiguous "middleman" minority position. This kind of middle tier status for individuals of mixed race is common in Latin America [@telles_race_2009] and @bonilla-silva_biracial_2004 has argued that such a system may be emerging in the contemporary US as well. In this case, many biracial groups may occupy a nebulous and ill-defined middleman or buffer class position in-between whites and darker-skinned non-whites, helping to cement a new kind of racial hierarchy.

## Differential Resources

Racial inequality is also at least partly explained by different distributions of material, cultural, and social resources across racial groups <!-- TODO CITATIONS HERE? -->. On average, socioeconomic background (e.g. parental income, wealth, and education) differs by race and such socioeconomic background is consequential for outcomes. Other less materially defined resources also vary across groups such as cultural styles, social networks, and neighborhood quality. When such characteristics are held constant or "controlled", the remaining differences between racial groups tend to be smaller, although such differences rarely disappear <!-- TODO WHAT CAN I CITE HERE -->. These differences in background resources reflect the accumulation of historical processes of discrimination <!-- CITE PAGER? -->. Thus, observed racial inequality reflects both contemporary processes of discrimination and bias and the intergenerational accumulation of disadvantage.

How do these differential distributions of resources affect biracial individuals? A naive argument would be that because biracial individuals have parents from two different racial groups, they will on average inherit parental resources roughly halfway in-between their two constituent groups [@chew_american_1989]. Prior research looking at biracial children in the Census data has found some evidence to support this argument [@chew_american_1989]. In such a case, the resources of biracial individuals would predict outcomes in-between the two constituent groups, barring other mechanisms.

However, the actual resource position of biracial individuals is likely far more complicated because individuals are not randomly selected into interracial unions. How this selection will affect a biracial child's outcomes for a given type of pairing are complex because it depends on the strength and direction of that selection on each partner, its consistency across different types of unions, and how strongly the selected characteristic predicts outcomes in the next generation.

Status exchange theory predicts that given an unequal racial hierarchy, individuals from a lower status racial group will be positively selected on socioeconomic characteristics into intermarriage with a higher status group, while individuals from a higher status racial group will be similarly negatively selected into the same marriage [@davis_intermarriage_1941; @merton_intermarriage_1941; @fu_racial_2001]. Using education as the socioeconomic parameter, prior research has shown strong evidence of this pattern in the case of Black male/White female marriages in the US, but weaker evidence for other types of unions [@gullickson_education_2006; @fu_racial_2001; @kalmijn_educational_2010; @hou_interracial_2011]. Less research has explored these patterns outside of marriage <!-- TODO cites? -->. In any case, because status exchange positively selects one parent and negatively selects the other partner, the two effects may somewhat cancel each other out in terms of the overall distribution of resources.

Regardless of the underlying complexity of this selection process, delineating its role as opposed to other mechanisms is key to understanding the overall process that sorts biracial individuals within the racial hierarchy. Although this approach is sometimes characterized as pitting race versus class and minimizing the importance of race, that is not case here. Instead, what this delineation allows us to do is understand how race affects the outcomes of biracial individuals in a multigenerational process to better pinpoint the generational locus of specific mechanisms.

<!-- TODO how can I integrate Cheng and Powell? -->

## Hybridity

The experience of hybridity itself may also positively and/or negatively affect outcomes for biracial individuals. The "marginal man" theory argues that biracial individuals will find it difficult to fit in because of the dissonance of an identity that crosses strong and salient racial boundaries [@park_human_1928; @stonequist_problem_1935]. This liminal identity conflict will then result in negative mental health experiences that could feed into other negative outcomes as well. Evidence for such negative mental health experiences, however, has been mixed [@udry_health_2003; @campbell_what_2006; @cheng_multiracial_2009; @bratter_does_2011].

Marginal man theory, developed in the first half of the Twentieth century, viewed culture in monolithic terms and thus to be trapped in between two static and monolithic cultures was seen as a difficult experience. Contemporary notions of culture emphasize its "toolkit" nature, arguing for a more flexible and fluid notion of culture <!-- TODO who to cite here? Swidler obviously, but also Dimaggio. Look at Dimaggio and Markus article -->. In this case, biracial individuals may benefit from an enlarged cultural toolkit that allows them to navigate a variety of diverse social contexts (i.e. code switch) <!-- TODO There are citations for this -->.

The complexity of the underlying mechanisms described above prevents simple and straightforward hypotheses regarding the position of biracial individuals. Different mechanisms may produce the same observed outcome and the weight of different mechanisms may also vary by groups as these groups may governed by different regimes of classification. However, comparing outcomes across biracial groups may help us better understand the likely mechanisms that drive biracial classification and to disregard others as implausible. Accounting for the effect of parental resources on outcomes to the best of our ability is also key to reaching any conclusions about the placement of biracial individuals within the racial hierarchy.

# Prior Work

Prior work on the outcomes for biracial individuals has been hampered by several methodological issues. With good reason, researchers have been reluctant to use data in which multiracial or biracial status is determined solely by the self-identification of respondents. We know that far fewer people identify as biracial than could feasibly do so based on their parents' races <!-- TODO cite morning? morning and Goldstein? who? --> and that various selection effects operate on the decision to identify as biracial among those who could [@roth_end_2005a; @davenport_role_2016]. Furthermore, data on changes in racial reporting over time suggests that other status markers like education and income may influence how people identify themselves [@saperstein_racial_2012; @saperstein_mulatto_2013]. For example, a person with one Black and one White parent who is routinely treated by others as Black and discriminated against as Black may be more likely to identify themselves as exclusively Black than a similar person of Black and White ancestry who experiences less consistent discrimination. Thus, any analysis of outcomes for biracial individuals based on self-identification must deal with a serious potential for reverse causation in the observed differences between biracial individuals and their constituent groups [@bratter_multiracial_2018].

Instead of self-identification, researchers have generally relied upon survey data in which the racial identification of a respondent's biological parents are provided in some form or another. @campbell_multiracial_2009 used data from the National Longitudinal Survey of Adolescent to Adult Health (Add Health) to examine high school GPA, advanced placement in math, and four-year college enrollment among biracial and monoracial individuals identified by co-resident biological parents' races. @herman_blackwhiteother_2009 used data on high school students in California and Wisconsin in which the students reported their parents' races to examine differences in GPA. Both of these studies find some support for an "in-between" status for many biracial groups, but are both hampered by a lack of statistical power due to small sample sizes. @kao_racial_1999 used data from the National Educational Longitudinal Study of 1988 to examine mathematics achievement scores and GPA for biracial Blacks and Asians, and finds that biracial Blacks have outcomes similar to monoracial Blacks, while biracial Asians have outcomes more similar to monoracial Whites. This study, however, is hampered by the fact that biracial students are identified by a discrepancy between their reported race and one parent's reported race, rather than by both biological parent's races.

The primary difficulty with the survey data used in the preceding studies is that the sample size for most biracial groups will be small and thus results tend to be imprecise and inconclusive. An alternative approach is to use large scale Census or American Community Survey (ACS) data to identify biracial children by the race of the biological parents in their household. This approach can generate much larger samples of biracial individuals, although it does exclude examination of children living in non-nuclear households. Furthermore, researchers then have a limited range of outcomes to explore because biracial individuals identified by this method are by definition children and the data sources include relatively little detailed information. This approach has been used previously to examine the likelihood of living in poverty [@bratter_multiracial_2013; @bratter_poverty_2013] and residential segregation [@ellis_agents_2012]. <!-- TODO briefly summarize findings --> While these findings provide valuable information of the context of biracial childrens' lived experience, but do not directly measure their own outcomes, but rather the experiences of their parents.

However, one important educational outcome is identifiable for children in Census and ACS data - grade retention. Grade retention is partially identifiable in these data sources and provides one of the earliest measurable forms of stratification that individuals face in their lives.

# Grade Retention as an Outcome

-   Why grade retention as an outcome?

    -   Grade retention is a relatively common practice. Discuss overall trends in grade retention and connection to educational reforms.
    -   Grade retention has negative effect on later outcomes, even controlling for cognitive differences (CITE WORK HERE).
    -   Clear inequalities by race in practices of grade retention as well as by SES (CITE WORK).
    -   In effect, grade retention may be for many individuals the first encounter with the wider stratification system that sorts individuals and helps to perpetuate inequality.

-   What I do and how it improves our understanding.

# Data and Methods

Data for this analysis come from the American Community Survey (ACS), an annual 1-in-100 survey of the US population, produced by the Census Bureau. To increase sample size for smaller populations of biracial respondents, I pool ACS data for a full decade from 2010-2019. While pooling data helps to increase sample size, it may generate biases if the likelihood of grade retention is changing over time in complex ways. I address this issue in the modeling techniques discussed below. All data were extracted from the Minnesota Population Center's Integrated Public Use Microdata Series [@ruggles_ipums_2020].

Ideally, I would restrict the sample only to children living in a household with two biological parents. However, the ACS only records the direct family relationship between the head of household andother members of the family. Since 2008, the ACS also distinguishes between biological children, adopted children, and stepchildren. Respondents can therefore be identified as biological children of another member of the household only if that member is the head of household or the biological child of the head of household. Determining whether the partner/spouse of that biological parent is the other biological parent is more difficult.

Prior work has used a variety of additional restrictions to limit the analysis to children who are more likely to be the biological children of both parents [@saenz_persistence_1995; @xie_racial_1997]. Some of those procedures rely upon an assumption that most parents would be married that was more tenable in historical data than contemporary data. Therefore, I do not explicitly restrict by marital status, I do restrict cases in two other ways that are consistent with prior work. First, I restrict the sample to cases where both parents were of a reasonable age at the birth of the child (aged 15-44 for mothers and aged 15-60 for fathers). Second, I restrict the sample to only those cases where the reported race and Hispanicity of the child is consistent with at least one of the parents.

The final analytical sample consists of children aged 5-20 who are currently enrolled in the K-12 school system and live in a household with two parents who are both likely to be biological parents. This necessary sample restriction will bias the family structure of the analytical sample relative to the total population of K-12 schoolchildren, limiting it exclusively to households with two parents present and over-representing the proportion of cases in which those parents are married. The result of this bias is well-known. Children living in two-parent nuclear households have greater educational success than other children as a result of material, social, and cultural resources (CITATIONS). Therefore, this sample restriction will upwardly bias estimates of grade retention. However, this bias is less problematic for the goal of this study, which is to estimate differences between groups. The sample restriction effectively eliminates differences in family structure across racial groups which may account for some observed differences in grade retention. However, the remaining differences observed here are likely to hold for the full population in the absence of strong racial differences in the effect of family structure on grade retention.

The race of children is calculated by a cross-tabulation of parent's races. A full cross-tabulation of the race/Hispanicity of each parent would produce far too many combined categories to be analytically useful. Therefore, I collapse each parent's race into the categories of White, Black, Asian, Latino, and Indigenous. These categories form the "ethnoracial pentagon" that is commonly used in popular practice and government tabulation to identify race in the US [@hollinger_postethnic_1995]. Parents are identified as Latino based on their response to the Hispanicity question, regardless of their response to the race question. Parents are identified as Indigenous if they responded as either American Indian/Alaska Native or as Pacific Islander. Because the goal of my analysis is to examine specifically the outcomes of "first-generation" biracial children, I exclude cases where at least one parent identified with multiple races.

The cross-tabulation of parent's races leads to ten distinct biracial categories. The total sample size for each of these biracial categories is shown in @tbl-descriptive, along with several descriptive statistics. The sample sizes are quite large for many of the groups, suggesting substantial statistical power in determining the relative outcomes for these groups. Even among the smallest group of Indigenous/Asian, the sample size is substantially greater than one would see in typical survey data.

```{r}
#| label: tbl-descriptive
#| tbl-cap: Descriptive survey-weighted statistics by race. Shading indicates biracial group. Groups ordered by sample size.
tbl_summary <- acs_design %>%
  group_by(race) %>%
  summarize(n=n(), 
            held_back=percent(as.numeric(survey_mean(below_exp_grade, 
                                                     vartype=NULL)), 
                              accuracy=0.01),
            mean_age=round(survey_mean(age, vartype=NULL), 1)) %>%
  arrange(desc(n)) %>%
  mutate(n=comma(n))

index_biracial <- which(str_detect(tbl_summary$race, "/"))

tbl_summary %>%
  kbl(booktabs = TRUE,
      linesep = NULL,
      align = c("l",rep("r", ncol(tbl_summary)-1)),
      col.names=c("race",
                  "n",
                  "clearly below expected grade", 
                  "mean age" )) %>%
  kable_styling() %>%
  row_spec(index_biracial, background = "#D3D3D3")
```

To measure grade retention, I compare the age of respondents to their current grade in school. Because the ACS lacks data on the exact birthday of respondents, it is not possible to exactly identify students who are below the expected grade level. A student of any given age in years can always be expected in two different grades. For example, a first grader is expected to be either six or seven years of age, while a second grader is expected to be either seven or eight years of age. Therefore, with information only on current grade and age in years, it is impossible to determine whether a seven year old in first grade is behind the expected grade level.

Instead of trying to identify exact grade retention, I instead identify cases in which students are clearly below expected grade. A student is clearly behind expected grade when the age of the student is greater than either of the possible ages for a student of that grade (e.g. a first grader who is eight years old). This measure will underestimate the total proportion of students that have been held back. However, the goal of this study is to examine differences across groups rather than total levels. The only factor determining this bias is whether the birthday of the student takes place earlier or later in the year, and is unlikely to vary substantially across groups.[^1]

[^1]: Biracial respondents are slightly younger than monoracial respondents in the sample as a whole because interracial unions are becoming more common over time HOW STRONG COULD THE BIAS BE HERE AND WHAT CAN I DO ABOUT IT?

==ACKNOWLEDGE POTENTIAL ISSUE WITH DROPOUT STUDENTS==

@tbl-descriptive shows the percent of each monoracial and biracial group that is clearly below the expected grade. My goal is to understand the degree of grade retention for each biracial group relative to its constituent monoracial groups (e.g. a Black/White student compared to White and Black students). Throughout this article, I use a visual approach to illustrate this placement.

As an example, I display the percent of students clearly behind grade for White, Black, and Black/White students in @fig-example. The point estimate and confidence interval for the Black/White group are shown in red, while the two monoracial groups are shown via grey dots and confidence bands. For all three cases, I use 83.4% confidence intervals, because overlap on such intervals indicates that the observed difference between the two measures in the sample is not statistically distinguishable from zero in the population at $p<.05$ (CITE).[^2] In this case, the lack of overlap between any of the confidence bands indicates that the Black/White probability of being clearly below expected grade is statistically distinguishable from both monoracial probabilities.

[^2]: The correspondence to $p<.05$ is only approximate due to differences in the two standard errors between estimates. However, this approximation will be extremely close in almost all cases.

```{r example-plot}
#| label: fig-example
#| fig-cap: Probability of being clearly below expected grade for biracial black/white respondents in comparison to their monoracial comparison groups. 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level.
g <-  coef_table %>%
  filter(mrace=="Black/White" & model=="Raw") %>%
  ggplot(aes(x=term, y=estimate, color=multiracial,
             ymin=estimate-1.386*se, ymax=estimate+1.386*se))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=estimate_mid-1.386*se_mid,
                ymax=estimate_mid+1.386*se_mid, fill=multiracial), 
            color=NA, alpha=0.2)+
  geom_linerange(data=subset(coef_table, multiracial & 
                               mrace=="Black/White" &
                               model=="Raw"))+
  geom_point()+
  coord_flip()+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="predicted probability of being behind expected grade")

g+annotate("text", 1, 0.036, hjust="left",
             label="Grey bands indicate 83.4%\nconfidence interval around\nmonoracial estimates")+
  annotate("text", 2, 0.0445, hjust="left",
             label="Red band indicate 83.4%\nconfidence interval around\nmidpoint estimate")+
  annotate("text", 2.5, 0.0365, hjust="left",
             label="Actual biracial estimate\nshown with 83.4%\nconfidence interval")
```

Additionally, I include a measure of the "halfway" point between the two racial groups.[^3] This halfway point allows me to determine where the biracial group falls relative to the expectation of being halfway between the two constituent monoracial groups. In this case, Black/White students are much closer to White students in their risk of grade retention, and have probabilities much lower than both Black students and the halfway expectation.

[^3]: The standard error of this halfway estimate used to construct the confidence interval is given by $\sqrt{(s_1^2+s_2^2)/2}$ where $s_1$ and $s_2$ are the standard errors of the estimates for the two monoracial groups.

@fig-example shows the raw differences between the three racial groups of interest. However, in practice, I want to estimate differences across groups while holding constant a variety of potentially confounding variables. To do this, I estimate a set of logit models that predict the likelihood of a student being clearly behind expected grade by race and a variety of other variables. I then construct figures similar to @fig-example by calculating from the model the average predicted probabilities (APP) for each racial group. APPs estimate the average probability of the outcome by some covariate across all cases, while holding constant all other variables. APPs are akin to average marginal effects (AME) and estimated in the same way via the delta method. Whereas AMEs estimate differences or slopes, APPs estimate the level for a given category.[^4]

[^4]: Output from the full models used to construct these APPs is available in the supplementary materials provided with this article.

I begin with a baseline model that adjusts for a variety of nuisance characteristics that may vary across groups and thus need to be controlled in all models. First, I include fixed effects for state of residence because states dictate educational policy and thus can differ substantially in the likelihood of grade retention. I cannot identify specific school districts, but I also include dummy variables indicating whether the student lived in a central city, suburban, or rural area.[^5]

[^5]: The location cannot be determined for all students and I therefore also include a fourth category of unknown in all models.

The probability of being clearly below expected grade also increases with the student's current grade, so I include fixed effects for the current grade of the student. Additionally, @fig-year-grade shows that the percent of students clearly behind grade level has declined over time, but this decline has been much more substantial at higher grade levels. For the elementary grades, there is no evidence of a decline at all. To account for this grade-specific decline in the models I include an interaction between a linear year term and current grade.[^6]

[^6]: Sensitivity analysis showed that this functional form was preferred by the Bayesian information criterion to models with no interaction and a model with interaction terms that treated year as a categorical variable. I also considered similar interactions between state and year, but none of the models with these interaction terms were preferred.

```{r}
#| label: fig-year-grade
#| fig-cap: Trends over time and grade in the percent of students clearly behind grade level. Lines are fit to each set of points by grade via lowess smoothing.

acs %>%
  group_by(current_grade, year) %>%
  summarize(below_grade=mean(below_exp_grade)) %>%
  ggplot(aes(x=year, y=below_grade, group=current_grade, color=current_grade))+
  #geom_line()+
  geom_smooth(se=FALSE)+
  geom_point()+
  scale_x_continuous(breaks=c(2010,2014,2018))+
  scale_y_continuous(labels=scales::percent)+
  labs(x="year", y="percent clearly behind grade level for age",
       color="current grade")+
  theme_bw()
```

The baseline model also includes the race of the respondent. I then add additional terms measuring material and cultural resources that may account for racial differences in a set of subsequent models. First, I include measures of nativity and English proficiency for both the respondent and each of their parents. Second, I include a categorical measure of highest degree earned for each parent. Finally, I include measures of family income (square rooted), home ownership, and whether the parents are married that capture other family resources. All standard errors reported here also account for the clustering of multiple respondents within the same household.

# Results

## Monoracial Differences

I begin by showing the relative risk of being below expected grade among monoracial respondents across models. @tbl-monoracial reports average marginal effects on the probability of grade retention for each monoracial minority group in comparison to White students. Understanding these differences and how they change across models helps me to understand the potential for how biracial respondents might be positioned.

Model 1 only controls for structural factors such as state of residence, student location, grade, and year. These results provide a baseline estimate of the differences across monoracial groups without controlling for differences in cultural and material resources across groups. The results show that Black, Indigenous, and Latino students all have substantially higher probability of grade retention than White students. Indigenous students have substantially higher risk than all other students, with a probability of grade retention that is 4.1% percentage points higher than White students. Asian students have the lowest probability of grade retention, and their risk is slightly lower than Whites.

The subsequent models control for a variety of additional variables. Model 2 includes controls for whether the student and each of their parents are native-born and speak English well. Model 3 controls for the highest degree received for each parent, and Model 4 controls for family income (square rooted), home ownership, and whether the student's parents are married.

In general Model 2 and Model 3 have substantial effects on the observed inequality across racial groups, while the additional variables from Model 4 have less impact. Controlling for the characteristics related to immigration in Model 2 cuts the gap between White and Latino students in half and increases the gap between Asian and White students. It has no impact on the gap between White and Black or White and Indigenous students because most of these students are native-born with native-born parents.

Controlling for education in Model 3 reduces the gap substantially for Black and Indigenous students and removes the gap entirely for Latino students. In fact, the results of both Models 3 and 4 suggest that, when holding constant cultural and family resources, Latino students are slightly less likely than White students to be held back.

SUMMARIZE THIS?

```{r monoracial-results}
#| results: asis
#| label: tbl-monoracial
#| tbl-cap: Average marginal differences in the probability of being behind grade by monoracial group, relative to a monoracial white student
texreg(map(results_monoracial[-1], convert_marg),
       digits=3,
       caption = "Average marginal differences in the probability of being clearly behind grade by monoracial group, relative to a monoracial white student",
       caption.above=TRUE,
       label="tbl-monoracial",  #TODO: This does not work
       float.pos="t",
       custom.gof.rows=
         list("State fixed effects" = c("Yes","Yes","Yes","Yes"),
              "Location fixed effects" = c("Yes","Yes","Yes","Yes"),
              "Grade fixed effects" = c("Yes","Yes","Yes","Yes"),
              "Year linear effects" = c("Yes","Yes","Yes","Yes"),
              "Year x grade effects"  = c("Yes","Yes","Yes","Yes"),
              "Nativity and language" = c("No","Yes","Yes","Yes"),
              "Parent's education" = c("No","No","Yes","Yes"),
              "Family resources" = c("No","No","No","Yes")))
```

## Resource Differences by Race

```{r}
#| label: tbl-resources
#| tbl-cap: Mean resources by racial group. All results are survey weighted. Results sorted by counterfactual probability of grade retention. Shading indicates biracial group.
index_biracial <- which(str_detect(tbl_resources$race, "/"))

tbl_resources %>%
  select(race, held_back_cf, fam_inc, own_home, m_ba, f_ba, f_born, par_fb,
         spk_eng_well, spk_eng_par) %>%
  mutate(held_back_cf=percent(held_back_cf/100, accuracy=0.1),
         own_home=percent(own_home/100, accuracy=0.1),
         m_ba=percent(m_ba/100, accuracy=0.1),
         f_ba=percent(f_ba/100, accuracy=0.1),
         f_born=percent(f_born/100, accuracy=0.1),
         par_fb=percent(par_fb/100, accuracy=0.1),
         spk_eng_well=percent(spk_eng_well/100, accuracy=0.1),
         spk_eng_par=percent(spk_eng_par/100, accuracy=0.1),
         fam_inc=dollar(fam_inc)) %>%
  kbl(booktabs = TRUE,
      linesep = NULL,
      align = c("l",rep("r", ncol(tbl_resources)-1)),
      col.names=c("race","held back, counterfactual","family income",
                  "own home","mother, 4-year college degree",
                  "father, 4-year college degree",
                  "foreign-born","either parent foreign-born",
                  "speak English well", 
                  "Both parents speak english well")) %>%
  kable_styling() %>%
  row_spec(index_biracial, background = "#D3D3D3") %>%
  column_spec(2:10, width = "1.5cm") %>%
  landscape()
```

|                   | Lower than both | Same as lower | Between | Same as higher | Higher than both |
|------------|:----------:|:----------:|:----------:|:----------:|:----------:|
| White/Asian       |        X        |               |         |                |                  |
| Black/Asian       |                 |   X (Asian)   |         |                |                  |
| Black/Latino      |                 |   X (Black)   |         |                |                  |
| White/Latino      |                 |   X (White)   |         |                |                  |
| Indigenous/Latino |                 |  X (Indig.)   |         |                |                  |
| Latino/Asian      |                 |   X (Asian)   |         |                |                  |
| White/Indigenous  |                 |               |    X    |                |                  |
| Indigenous/Asian  |                 |               |    X    |                |                  |
| Black/Indigenous  |                 |               |    X    |                |                  |
| Black/White       |                 |               |         |   X (Black)    |                  |

: Relative placement for biracial groups in comparison to monoracial constituent groups of counterfactual predicted probability of being held back based due to resources.

## Biracial Placement

I now turn to the results for biracial respondents. Due to the number of comparisons being made, I split these results into two separate figures. In @fig-black-models, I show the placement of each biracial group involving one black parent in comparison to their two monoracial constituent groups. To allow comparison across models, I compare the results from the baseline model (Model 1 in @tbl-monoracial) and the full model that accounts for all cultural and family resources (Model 4 in @tbl-monoracial).

```{r fig-black-models}
#| label: fig-black-models
#| fig-pos: p
#| fig-cap: Probability of being behind grade for biracial respondents with one black parent, in comparison to their monoracial comparison groups. 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level, Baseline model includes year, grade, location, and state fixed effects. Full models include control variables for nativity, English proficiency, income, education, home ownership, and marital status of parents.
#| fig-height: 7.5
coef_table %>%
  filter(model=="Baseline model" | model=="Full model") %>%
  filter(str_detect(mrace, "Black")) %>%
  ggplot(aes(x=term, y=estimate, color=multiracial,
             ymin=estimate-1.386*se, ymax=estimate+1.386*se))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=estimate_mid-1.386*se_mid,
                ymax=estimate_mid+1.386*se_mid, fill=multiracial), 
            color=NA, alpha=0.2)+
  geom_linerange(data=subset(coef_table, multiracial 
                             & (model=="Baseline model" | 
                                  model=="Full model") & 
                               str_detect(mrace, "Black")))+
  geom_point()+
  coord_flip()+
  facet_grid(mrace~model, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average predicted probability of being clearly behind expected grade")
```

I focus first on biracial groups with one Black parent because of a stronger historical expectation that these students will be identified as Black by the "one-drop rule" and as a result will have similar outcomes to Black students. In three cases, this hypothesis can be soundly rejected. For Black/White, Black/Latino, and Black/Asian students, The probability of being clearly below expected grade is substantially lower than that for monoracial black students and the difference is statistically distinguishable from zero. Only in the case of Black/Indigenous students, do I observe outcomes that are indistinguishable from those of Black students. However, in this case,monoracial Black students have lower risk of grade retention than monoracial Indigenous students. Thus, In all four cases, part-Black biracial students have a risk of grade retention that is closer to the lower risk monoracial group.

The results also suggest that each of these biracial groups has a risk of grade retention lower than the halfway point, although this hypothesis cannot be rejected completely for the Black/Asian and Black/Indigenous groups, due to a higher degree of uncertainty for these smaller groups.

Controlling for resource variables has some effect on the placement of Black/White and Black/Latino biracial respondents. In the case of Black/White biracial respondents, Black/White students initially have a risk of grade retention slightly higher than and statistically distinguishable from White students. After controlling for cultural and family resources, however, Black/White respondents have a risk of grade retention that is slightly lower and statistically distinguishable from White students. This change reflects the relatively low level of parental education and income among Black/White students, which is much closer to Black students than White students, as shown on @tbl-descriptive.

An opposite pattern is observed for Black/Latino respondents. Black/Latino respondents initially have a risk of grade retention substantially lower than both Black and Latino students, who have similar risks. However, after controlling for cultural and family resources, Black/Latino students have a risk of grade retention that is indistinguishable from the grade retention risk of Latino students and substantially lower than Black students. The change largely reflects the fact that Black/Latino students have greater cultural resources than Latino students and once this advantage is accounted for, they have a similar risk.

Altogether, the results here strongly suggest that the risk of grade retention faced by these biracial students with one black parent is generally similar to that of their constituent monoracial group with the lower risk of grade retention. This finding is inconsistent both with the expectation of the one-drop rule and the "in-between" hypothesis.

@fig-other-models shows the results for the remaining biracial groups. In general, the results here are consistent with those for part-Black biracial respondents, with some important caveats discussed below. The point estimates indicate that the risk of each biracial group is either quite close to the monoracial group with the lower risk or in-between this group's risk and the halfway point.

```{r fig-other-models}
#| label: fig-other-models
#| fig-pos: p
#| fig-cap: Probability of being behind grade for non-black biracial respondents, in comparison to their monoracial comparison groups. 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level, Baseline model includes year, grade, location, and state fixed effects. Full models include control variables for nativity, English proficiency, income, education, home ownership, and marital status of parents.
#| fig-height: 7.5
coef_table %>%
  filter(model=="Baseline model" | model=="Full model") %>%
  filter(!str_detect(mrace, "Black")) %>%
  ggplot(aes(x=term, y=estimate, color=multiracial,
             ymin=estimate-1.386*se, ymax=estimate+1.386*se))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=estimate_mid-1.386*se_mid,
                ymax=estimate_mid+1.386*se_mid, fill=multiracial), 
            color=NA, alpha=0.2)+
  geom_linerange(data=subset(coef_table, multiracial 
                             & (model=="Baseline model" | 
                                  model=="Full model") & 
                               !str_detect(mrace, "Black")))+
  geom_point()+
  coord_flip()+
  facet_grid(mrace~model, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average predicted probability of being clearly behind expected grade")
```

For all groups, the full models generally indicate that each biracial group has a risk of grade retention similar to the risk of the lower monoracial constituent group or has risk higher than this group but less than the halfway point. There are two exceptions to this general pattern. First, for the Indigenous/Latino group, I cannot rule out the possibility that this group has risk roughly halfway between the two constituent groups, although the point estimate favors a result closer to the lower Latino group. Second, Latino/Asian students are the only case where the risk for the biracial group is higher than either monoracial group. However, the confidence bands suggest that the risk for the Latino/Asian group cannot be differentiated from the rates for Latinos, the higher of the two monoracial constituent groups. This result marks the one substantial deviation from the general pattern. It only emerges in the models that control for cultural and family resources because Latino/Asian students come from higher income and more educated households than Latino students, on average.

Aside from the effect on Latino/Asian students, controlling for cultural and family resources has moderate effects on the placement of the remaining biracial groups. The only case with a substantial shift is that for White/Asian students who initially appear to have lower risk than either White or Asian students. After controlling for these resources, the risks of White/Asian students are indistinguishable from the lower risk of Asian students. This is due to the fact that White/Asian students come from the most advantaged households of any group in the dataset.

Any results for part-Latino students are complicated by the substantial change in the monoracial Latino risk across models. For example, while Latinos have substantially higher risk of grade retention than Whites in the baseline model, there is only a very small difference favoring Latinos in the full model. Nonetheless, even in this case, we see that White/Latinos have risk closer to those of the lower Latino group, but distinguishable from both groups.

# Conclusions

-   Big claim: The results are consistent with a new "regime of hyperdescent" operating across all racial groups. This contributes to a muddying of racial distinctiveness via hybridity, while at the same time isolating monoracial members of the most stigmatized racial groups, consistent with arguments like Bonilla-Silva's Latin Americanization.

# Acknowledgments {.appendix}

I am grateful to...

# References {.appendix}