### 2022-09-09 products/paper/main.qmd
Update figure code for paper

---
title: Grade Retention and Mixed Race
author: 
  - Aaron Gullickson, *University of Oregon*
abstract: This is a test abstract
thanks: The corresponding author can be reached at aarong@uoregon.edu. All code and data for this project are available at ...
format:
  pdf:
    documentclass: scrartcl
    papersize: letter
    fig-height: 4
    fig-pos: t
    keep-tex: false
  docx:
    toc: false
bibliography: ../project.bib
execute: 
  eval: true
  echo: false
  warning: false
  error: false
editor: source
---

```{r setup}
#| include: false
library(here)
source(here("analysis","check_packages.R"))
source(here("analysis","useful_functions.R"))
color_choices <- c("#003f5c","#58508d","#bc5090","#ff6361","#000000","#ffa600")
```

```{r load-data}
load(here("analysis","output","acs.RData"))
load(here("analysis","output","coef_table.RData"))
load(here("analysis","output","monoracial_results.RData"))
load(here("analysis","output","tbl_resources.RData"))
# change names of models for display
coef_table <- coef_table %>%
  mutate(model=factor(model,
                      levels=c("crude","fixed effects","cultural resources",
                               "parent resources"),
                      labels=c("Raw","Baseline model","Cultural model",
                               "Full model")))
ci_df <- ci_df %>%
  mutate(model=factor(model,
                      levels=c("crude","fixed effects","cultural resources",
                               "parent resources"),
                      labels=c("Raw","Baseline model","Cultural model",
                               "Full model")))
# create design object
acs_design <- acs %>%
  as_survey_design(ids = cluster, weights = perwt)
```

# Introduction

The experiences of mixed race individuals will play an important role in the future of race in the US. A rise in interracial marriage and coupling since the 1960s has resulted in a "biracial baby boom" of individuals whose parents cross racial lines [@root_racially_1992]. This demographic phenomenon may have a potentially transformative effect on racial boundaries [@hochschild_creating_2012; @alba_great_2020]. Although race mixing has always been a feature of US history [@morning_who_2000; @gullickson_choosing_2011], this biracial boom is notable for its demographic size and for occurring in a post-Civil Rights era in which historical norms governing how to handle racial mixture may be shifting. Understanding the transformative potential of this biracial baby boom is contingent on developing a better understanding of how these biracial individuals identify their own race and how they are sorted into a prevailing racial hierarchy that reinforces a system of racial inequality.

Extensive scholarship documents the identification and classification decisions of multiracial individuals [@xie_racial_1997; @roth_end_2005a; @brunsma_interracial_2005; @herman_forced_2004; @holloway_place_2009] <!-- TODO MORE HERE -->. This research has generally identified a surprising degree of fluidity and choice in classification decisions, in the context of historical norms of racial classification <!-- TODO CHECK -->. On the other hand, only a smattering of research has documented how the life chances of biracial individuals compare to their monoracial peers [@kao_racial_1999; @campbell_multiracial_2009; @herman_blackwhiteother_2009]. The lack of work on this topic, despite its acknowledged relevance, results from two important limitations in most available data sources.

First, using direct identification of biracial respondents on data sources such as the Census or large scale surveys is problematic. Research on the identification and classification of biracial individuals indicates that individuals whose parents belong to different races will not always or consistently identify with both races [@doyle_are_2007; @liebler_america_2017; @gullickson_racial_2019]. Furthermore, social status and social mobility has been shown to predict racial identification generally [@saperstein_racial_2012; @saperstein_mulatto_2013] and social status has been shown to be related to the identification choices of biracial individuals in particular [@roth_end_2005a; @davenport_role_2016]. Thus, differences in outcome may in some part produce differences in identification rather than the other way around when examining self-identified biracial individuals.

This problem of reverse causation can be largely addressed when researchers have access to the racial identification of a respondent's biological parents. While such information is available on some large scale longitudinal surveys, researchers then run into a second problem of sample size. The samples of biracial respondents in such cases are generally far too small to make reliable statistical inferences. This limitation is particularly severe for biracial individuals with two non-white parents.

When children reside with both biological parents, biracial individuals can be identified in large scale data sources such as the Census. Researchers have long utilized the ability to identify such biracial children to examine the classification choices made on behalf of these children [@xie_racial_1997; @roth_end_2005a]. Because of their age, it is more difficult to analyze the outcomes of such children due to the simple fact that they haven't had many measurable life outcomes yet.

However, one important outcome, grade retention, is measurable and important even for young children. Being held back in school is a relatively common occurrence that has consequences for later educational transitions and is disproportionately applied by race <!-- TODO CITATIONS -->. Grade retention is arguably one of the earliest stratifiers of life chances that an individual will experience and within it we can clearly see the imprint of the racial hierarchy. In this article, I use data from the American Community Survey to examine the likelihood of being held back in school for biracial children in comparison to their monoracial peers. Pooling data from 2010 to 2019, I am able to draw a large enough sample to make reliable statistical inferences for most biracial groups, including those with two non-white parents.

The results provide a window into the placement of biracial children within an existing system of racial inequality in our educational systems. <!-- TODO  PITHY SUMMARY -->

# Where do Biracials Fit?

How might we expect the outcomes for biracial individuals do differ from their monoracial peers? In every case, the relevant comparison is how a biracial person fares relative to the two constituent groups with which their parents identify. For example, in the case of an individual with one Black and one White parent, we would compare the average outcomes for such individuals to the average outcomes for individuals with two White and two Black parents. The outcome for the Black/White group may be similar to one of the monoracial groups, it may be somewhere in between the outcomes for the two monoracial groups, or it may be higher or lower than either of the monoracial groups.

To better understand where biracial individuals might fit, it is necessary to understand what is producing a contemporary system of racial inequality overall. Here, I distinguish three mechanisms that may help us understand the biracial position within this racial hierarchy. First, I consider arguments for the racial hierarchy that focus on the effects of racial discrimination at both an individual and systemic level. Second, I consider arguments regarding different distributions of resources across racial groups that produce different average outcomes. Third, I consider arguments regarding how the nature of hybridity itself may affect outcomes.

## Discrimination and Classification

Discrimination and bias by race play an important role in overall levels of racial inequality. Discrimination need not be consciously recognized by actors, but may instead reflect cognitive biases and schema that individuals unknowningly rely upon and reinforce in social interactions <!-- TODO citations, may be Pager, Brubaker? -->. Such discrimination will be most consequential when performed by important gatekeepers of opportunity such as teachers, health care workers, and employers. Discrimination need not just occur on an individual basis, but can also be systemic in nature when biases are built into organizations and institutional systems <!-- TODO citations, Tilly? Bonilla-Silva, Pager again? -->. In our current "colorblind" regime, such systems may be formally race neutral but produce disparate impacts by race <!-- TODO CITE Bonilla-Silva? -->.

The experiences of biracial individuals within such a system of discrimination depend almost entirely on how they are classified by others. To the extent that biracial individuals are classified as more like one of their monoracial constituent groups than the other, they will receive the same treatment and reap similar rewards or suffer similar penalties.

The strongest historical precedent for understanding such classification in the US is the "one drop rule" norm concerning Black ancestry. According to the one-drop rule, individuals of known Black ancestry are considered exclusively Black, regardless of other ancestries[@davis_who_1991]. @davis_who_1991 generalizes this practice into the concept of *hypodescent* in which individuals with mixed racial ancestry will be classified as members of the lower status group. By comparison a pattern of *hyperdescent* implies that individuals of mixed racial ancestry will be classified as members of the higher status group. This pattern of hyperdescent has historically governed the experiences of individuals with American Indian ancestry <!-- TODO CITE -->.

Although the hypodescent/hyperdescent paradigm is often used to theoretically frame the possible contemporary experiences of multiracial individuals <!-- TODO citations -->, there are reasons to be cautious about the generalizability of this paradigm to both the contemporary and historical situation. Hypodescent and hyperdescent are observed in two specific historical cases involving Black ancestry and American Indian ancestry, respectively. As @iverson_regimes_2022 note, these two cases could just as parsimoniously be explained by a "dominance" model which ranks ancestries by their tendencies to be supercessive or recessive in determining identification and classification. In this framework, within the historical "racial triangle" of American colonialism, Black ancestry is supercessive, Indigenous ancestry is recessive, and White ancestry falls somewhere in between.

The development of these particular norms may also reflect the historical regime in which they were developed [@gullickson_choosing_2011; @iverson_regimes_2022]. In this case, neither the hypodescent/hyperdescent or dominance paradigm may help us to fully understand the experiences of those of Latino and Asian ancestry due to more recent processes of racialization. Second, the historical norms for all groups may also be breaking up or transforming in the post-Civil Rights era <!-- TODO CITE -->.

If outcomes for biracial individuals all closer to one of their constituent groups, then we may have stronger evidence about such norms of classification. However, what if those outcomes fall in-between the two groups? How might classification and discrimination produce such an in-between status? There are several possibilities here.

First, racial ambiguity in appearance may lead to more inconsistent classification into single race categories across observers. Even if norms consistently indicate a certain type of classification, correct identification by observers is still a probabilistic phenomenon in all social interactions and likely to be more heterogeneous for biracial individuals. As a result, biracial individuals will experience a lower sum total of discrimination and bias affecting the lower status group in the case of hypodescent and will not reap the same rewards as a member of the higher status group in the case of hyperdescent.

Second, even when "correctly" classified, greater physical ambiguity may allow biracial individuals to avoid the worst instances of bias and discrimination because they do not conform to the ideal type for the group facing the most discrimination. In other words, biracial people's experience of discrimination may be "softened" because their ambiguous physical appearance provokes less racial antipathy. The US in fact already has a long history of this kind of "softening." Because of the one-drop rule, Black individuals vary substantially in skin tone and those individuals with a lighter skin tone have better outcomes [@keith_skin_1991; @hughes_significance_1990; @monk_skin_2014] and are viewed more favorably by both Blacks and Whites [@maddox_cognitive_2002a]. Thus, light skin moderates the stigma of Blackness even among monoracial individuals.

Third, it is also possible that biracial individuals may be classified into an ambiguous "middleman" minority position. This kind of middle tier status for individuals of mixed race is common in Latin America [@telles_race_2009] and @bonilla-silva_biracial_2004 has argued that such a system may be emerging in the contemporary US as well. In this case, many biracial groups may occupy a nebulous and ill-defined middleman or buffer class position in-between whites and darker-skinned non-whites, helping to cement a new kind of racial hierarchy.

## Differential Resources

Racial inequality is also at least partly explained by different distributions of material, cultural, and social resources across racial groups <!-- TODO CITATIONS HERE? -->. On average, socioeconomic background (e.g. parental income, wealth, and education) differs by race and such socioeconomic background is consequential for outcomes. Other less materially defined resources also vary across groups such as cultural styles, social networks, and neighborhood quality. When such characteristics are held constant or "controlled", the remaining differences between racial groups tend to be smaller, although such differences rarely disappear <!-- TODO WHAT CAN I CITE HERE -->. These differences in background resources reflect the accumulation of historical processes of discrimination <!-- CITE PAGER? -->. Thus, observed racial inequality reflects both contemporary processes of discrimination and bias and the intergenerational accumulation of disadvantage.

How do these differential distributions of resources affect biracial individuals? A naive argument would be that because biracial individuals have parents from two different racial groups, they will on average inherit parental resources roughly halfway in-between their two constituent groups [@chew_american_1989]. Prior research looking at biracial children in the Census data has found some evidence to support this argument [@chew_american_1989]. In such a case, the resources of biracial individuals would predict outcomes in-between the two constituent groups, barring other mechanisms.

However, the actual resource position of biracial individuals is likely far more complicated because individuals are not randomly selected into interracial unions. How this selection will affect a biracial child's outcomes for a given type of pairing are complex because it depends on the strength and direction of that selection on each partner, its consistency across different types of unions, and how strongly the selected characteristic predicts outcomes in the next generation.

Status exchange theory predicts that given an unequal racial hierarchy, individuals from a lower status racial group will be positively selected on socioeconomic characteristics into intermarriage with a higher status group, while individuals from a higher status racial group will be similarly negatively selected into the same marriage [@davis_intermarriage_1941; @merton_intermarriage_1941; @fu_racial_2001]. Using education as the socioeconomic parameter, prior research has shown strong evidence of this pattern in the case of Black male/White female marriages in the US, but weaker evidence for other types of unions [@gullickson_education_2006; @fu_racial_2001; @kalmijn_educational_2010; @hou_interracial_2011]. Less research has explored these patterns outside of marriage <!-- TODO cites? -->. In any case, because status exchange positively selects one parent and negatively selects the other partner, the two effects may somewhat cancel each other out in terms of the overall distribution of resources.

Regardless of the underlying complexity of this selection process, delineating its role as opposed to other mechanisms is key to understanding the overall process that sorts biracial individuals within the racial hierarchy. Although this approach is sometimes characterized as pitting race versus class and minimizing the importance of race, that is not case here. Instead, what this delineation allows us to do is understand how race affects the outcomes of biracial individuals in a multigenerational process to better pinpoint the generational locus of specific mechanisms.

<!-- TODO how can I integrate Cheng and Powell? -->

## Hybridity

The experience of hybridity itself may also positively and/or negatively affect outcomes for biracial individuals. The "marginal man" theory argues that biracial individuals will find it difficult to fit in because of the dissonance of an identity that crosses strong and salient racial boundaries [@park_human_1928; @stonequist_problem_1935]. This liminal identity conflict will then result in negative mental health experiences that could feed into other negative outcomes as well. Evidence for such negative mental health experiences, however, has been mixed [@udry_health_2003; @campbell_what_2006; @cheng_multiracial_2009; @bratter_does_2011].

Marginal man theory, developed in the first half of the Twentieth century, viewed culture in monolithic terms and thus to be trapped in between two static and monolithic cultures was seen as a difficult experience. Contemporary notions of culture emphasize its "toolkit" nature, arguing for a more flexible and fluid notion of culture <!-- TODO who to cite here? Swidler obviously, but also Dimaggio. Look at Dimaggio and Markus article -->. In this case, biracial individuals may benefit from an enlarged cultural toolkit that allows them to navigate a variety of diverse social contexts (i.e. code switch) <!-- TODO There are citations for this -->.

The complexity of the underlying mechanisms described above prevents simple and straightforward hypotheses regarding the position of biracial individuals. Different mechanisms may produce the same observed outcome and the weight of different mechanisms may also vary by groups as these groups may governed by different regimes of classification. However, comparing outcomes across biracial groups may help us better understand the likely mechanisms that drive biracial classification and to disregard others as implausible. Accounting for the effect of parental resources on outcomes to the best of our ability is also key to reaching any conclusions about the placement of biracial individuals within the racial hierarchy.

# Prior Work

Prior work on the outcomes for biracial individuals has been hampered by several methodological issues. With good reason, researchers have been reluctant to use data in which multiracial or biracial status is determined solely by the self-identification of respondents. We know that far fewer people identify as biracial than could feasibly do so based on their parents' races <!-- TODO cite morning? morning and Goldstein? who? --> and that various selection effects operate on the decision to identify as biracial among those who could [@roth_end_2005a; @davenport_role_2016]. Furthermore, data on changes in racial reporting over time suggests that other status markers like education and income may influence how people identify themselves [@saperstein_racial_2012; @saperstein_mulatto_2013]. For example, a person with one Black and one White parent who is routinely treated by others as Black and discriminated against as Black may be more likely to identify themselves as exclusively Black than a similar person of Black and White ancestry who experiences less consistent discrimination. Thus, any analysis of outcomes for biracial individuals based on self-identification must deal with a serious potential for reverse causation in the observed differences between biracial individuals and their constituent groups [@bratter_multiracial_2018].

Instead of self-identification, researchers have generally relied upon survey data in which the racial identification of a respondent's biological parents are provided in some form or another. @campbell_multiracial_2009 used data from the National Longitudinal Survey of Adolescent to Adult Health (Add Health) to examine high school GPA, advanced placement in math, and four-year college enrollment among biracial and monoracial individuals identified by co-resident biological parents' races. @herman_blackwhiteother_2009 used data on high school students in California and Wisconsin in which the students reported their parents' races to examine differences in GPA. Both of these studies find some support for an "in-between" status for many biracial groups, but are both hampered by a lack of statistical power due to small sample sizes. @kao_racial_1999 used data from the National Educational Longitudinal Study of 1988 to examine mathematics achievement scores and GPA for biracial Blacks and Asians, and finds that biracial Blacks have outcomes similar to monoracial Blacks, while biracial Asians have outcomes more similar to monoracial Whites. This study, however, is hampered by the fact that biracial students are identified by a discrepancy between their reported race and one parent's reported race, rather than by both biological parent's races.

The primary difficulty with the survey data used in the preceding studies is that the sample size for most biracial groups will be small and thus results tend to be imprecise and inconclusive. An alternative approach is to use large scale Census or American Community Survey (ACS) data to identify biracial children by the race of the biological parents in their household. This approach can generate much larger samples of biracial individuals, although it does exclude examination of children living in non-nuclear households. Furthermore, researchers then have a limited range of outcomes to explore because biracial individuals identified by this method are by definition children and the data sources include relatively little detailed information. This approach has been used previously to examine the likelihood of living in poverty [@bratter_multiracial_2013; @bratter_poverty_2013] and residential segregation [@ellis_agents_2012]. <!-- TODO briefly summarize findings --> While these findings provide valuable information of the context of biracial childrens' lived experience, but do not directly measure their own outcomes, but rather the experiences of their parents.

However, one important educational outcome is identifiable for children in Census and ACS data - grade retention. Grade retention is partially identifiable in these data sources and provides one of the earliest measurable forms of stratification that individuals face in their lives.

# Grade Retention as an Outcome

-   Define what grade retention is and the era of "social promotion." Recent work suggests a decline in grade retention. Can cite Warren and Saliba on the trends.
-   Grade retention is consequential for later educational outcomes. CITE the evidence on this although, Warren and Saliba are also very iff-y on this. The Andrew article on "scarring" effects of grade retention is also a good one here as is the Hauser et. al. 2001 piece. Furthermore, little evidence on the positive effects of grade retention (again cite Hauser 2001). Grade retention also varies by race and socioeconomic background and is thus can be seen as a mechanism for the reproduction of class and racial inequality. In fact, because grade retention can happen quite early in the K-12 system, for many children, it may be the first direct encounter with a "sorting" mechanism for inequality.
-   There are many challenges to measuring grade retention in existing data sources. One of the most common techniques is to flag students who are behind the expected grade for their age. This technique is imperfect because
    -   it captures a cumulative prevalence measure of grade retention rather than incidence.
    -   it cannot distinguish cases in which children were delayed in beginning school ("redshirted") from grade retention.
    -   Students can reasonable be in two modal grades based on a given calendar year of age. Without detailed information on birthdates and survey timing, overall grade retention will be underestimated. For example, a child who is reported as 9 years old and in third grade could either be in the expected grade or could have been held back, depending on their birthday. A child who is 10 years old, however, is clearly behind expected grade.
-   Unfortunately, the only measure of grade retention available in the American Community Survey is based on students being clearly behind expected grade. The CPS has some advantages for better measuring grade retention, but lacks the sample size to precisely estimate parameters for biracial populations.
-   While being clearly behind expected grade is a problematic measure of grade retention for researchers seeking to better understand the process of grade retention itself, that is not my goal here. Rather, I am using this measure as an early and accessible measure of racial stratification for young children and adolescents. I can thus use it as a window to understand biracial students' placement within a racial hierarchy. This approach is similar to that employed by Rosenfeld (CITE) to examine the differential experiences of children from straight and LGBT families. From this perspective, the major potential problem is that the overall measure underestimates grade retention. However, because this underestimation is largely a function of birthdate, it is unlikely to bias the results presented here.
-   There is a question of how much racial inequality is present in grade retention rates. Hauser et. al. 2001 suggest that while Black and Latino students are more likely to have been held back, those differences are largely explained by differences in STUFF. However, other more recent research finds at least moderate racial differences, particularly between Black an White students, even after controlling for a variety of other variables (cite Warren and Sariba, Rosenfeld, others?). Therefore, understanding the effect of socioeconomic and other resources on the differences between monoracial groups is also key to understanding the larger research question, just as it is for biracial students.

# Data and Methods

Data for this analysis come from the American Community Survey (ACS), an annual 1-in-100 survey of the US population, conducted by the Census Bureau. To increase sample size for smaller populations of biracial respondents, I pool ACS data for a full decade from 2010-2019. While pooling data helps to increase sample size, it may generate biases if the likelihood of grade retention is changing over time in complex ways. I address this issue in the modeling techniques discussed below. All data were extracted from the Minnesota Population Center's Integrated Public Use Microdata Series [@ruggles_ipums_2020].

Ideally, I would restrict the sample only to children living in a household with two biological parents. However, the ACS only records the direct family relationship between the head of household and other members of the family. Since 2008, the ACS also distinguishes between biological children, adopted children, and stepchildren. Respondents can therefore be identified as biological children of another member of the household only if that member is the head of household or the biological child of the head of household. Determining whether the partner/spouse of that biological parent is the other biological parent is more difficult.

Prior work has used a variety of additional restrictions to limit the analysis to children who are more likely to be the biological children of both parents [@saenz_persistence_1995; @xie_racial_1997]. Some of those procedures rely upon an assumption that most parents would be married that was more tenable in historical than contemporary data. I do not explicitly restrict by marital status, but I do restrict cases in two other ways that are consistent with prior work. First, I restrict the sample to cases where both parents were of a reasonable age at the birth of the child (aged 15-44 for mothers and aged 15-60 for fathers). Second, I restrict the sample to only those cases where the reported race and Hispanicity of the child is inclusive of at least one of the parents' races.

The final analytical sample consists of children aged 5-20 who are currently enrolled in the K-12 school system and live in a household with two parents who are both likely to be biological parents. This necessary sample restriction will bias the family structure of the analytical sample relative to the total population of K-12 schoolchildren, limiting it exclusively to households with two parents present. The result of this bias is well-known. Children living in two-parent nuclear households have greater educational success than other children as a result of material, social, and cultural resources [@biblarz_family_1999]. Therefore, this sample restriction will bias overall estimates of grade retention downward. However, this bias is less problematic for the goal of this study, which is to estimate differences between groups. The sample restriction effectively eliminates differences in family structure across racial groups which may account for some observed differences in grade retention. However, the remaining differences observed here are likely to hold for the full population in the absence of strong racial differences in the effect of family structure on grade retention.

The race of children is calculated by a cross-tabulation of parent's races. A full cross-tabulation of the race/Hispanicity of each parent would produce far too many combined categories to be analytically useful. Therefore, I collapse each parent's race into the categories of White, Black, Asian, Latino, and Indigenous. These categories form the "ethnoracial pentagon" that is commonly used in popular practice and government tabulation to identify race in the US [@hollinger_postethnic_1995]. Parents are identified as Latino based on their response to the Hispanicity question, regardless of their response to the race question. Parents are identified as Indigenous if they responded as either American Indian/Alaska Native or as Pacific Islander. Because the goal of my analysis is to examine specifically the outcomes of "first-generation" biracial children, I exclude cases where at least one parent identified with multiple races.

The cross-tabulation of parent's races leads to ten distinct biracial categories. The total sample size for each of these biracial categories is shown in @tbl-descriptive, along with the sample size of the monoracial comparison groups. The sample sizes are quite large for most groups and far in excess of comparable categories in other available survey data.

```{r}
#| label: tbl-descriptive
#| tbl-cap: Descriptive survey-weighted statistics by race. Shading indicates biracial group. Groups ordered by sample size.
tbl_summary <- acs_design %>%
  group_by(race) %>%
  summarize(n=n(), 
            held_back=percent(as.numeric(survey_mean(below_exp_grade, 
                                                     vartype=NULL)), 
                              accuracy=0.01),
            mean_age=round(survey_mean(age, vartype=NULL), 1)) %>%
  arrange(desc(n)) %>%
  mutate(n=comma(n))

index_biracial <- which(str_detect(tbl_summary$race, "/"))

tbl_summary %>%
  kbl(booktabs = TRUE,
      linesep = NULL,
      align = c("l",rep("r", ncol(tbl_summary)-1)),
      col.names=c("Race",
                  "Sample size",
                  "Clearly behind expected grade", 
                  "Mean age" )) %>%
  kable_styling() %>%
  row_spec(index_biracial, background = "#D3D3D3")
```

Measuring grade retention exactly is not possible in the ACS data. Without data on date of birth and survey timing, a student of any given age who has not been held back could reasonably be in two different grades. For example, a first grader is expected to be either six or seven years of age, while a second grader is expected to be either seven or eight years of age. Therefore, with information only on current grade and age in years, it is impossible to determine whether a seven year old in first grade is behind the expected grade level or in the expected grade.

Instead of trying to identify exact grade retention, I instead identify cases in which students are clearly behind expected grade. A student is clearly behind expected grade when the age of the student is greater than either of the possible ages for a student of that grade (e.g. a first grader who is eight years old). This indirect measure of grade retention has been frequently used in previous studies of grade retention outcomes [@bianchi_children_1984; @frederick_have_2008; @rosenfeld_nontraditional_2010]. This measure will underestimate the total proportion of students that have been held back and may also suffer from other biases [@warren_patterns_2014]. However, the goal of this study is to examine differences across groups rather than total levels. The primary factor determining bias is whether the birthday of the student takes place earlier or later in the year, which is unlikely to vary substantially across groups.[^1]

[^1]: Biracial respondents are slightly younger than monoracial respondents in the sample as a whole because interracial unions are becoming more common over time HOW STRONG COULD THE BIAS BE HERE AND WHAT CAN I DO ABOUT IT?

@tbl-descriptive shows the percent of each monoracial and biracial group that is clearly behind expected grade. My goal is to understand the degree of grade retention for each biracial group relative to its constituent monoracial groups (e.g. a Black/White student compared to White and Black students). Throughout this article, I use a visual approach to illustrate this placement.

As an example, I display the percent of students clearly behind grade for White, Black, and Black/White students in @fig-example. The point estimate and confidence interval for the Black/White group are shown in red, while the two monoracial groups are shown via grey dots and confidence bands. For all three cases, I use 83.4% confidence intervals, because overlap on such intervals indicates that the observed difference between the two measures in the sample is not statistically distinguishable from zero in the population at $p<.05$ [@knol_mis_2011].[^2] In this case, the lack of overlap between any of the confidence bands indicates that the Black/White probability of being clearly behind expected grade is statistically distinguishable from both monoracial probabilities.

[^2]: The correspondence to $p<.05$ is only approximate due to differences in the standard errors between groups. Calculating more precise intervals requires that those intervals vary across each pairwise comparison which does not work for the visualizations here. The approximation will be very close in all cases used here.

```{r example-plot}
#| label: fig-example
#| fig-cap: Probability of being clearly below expected grade for biracial black/white respondents in comparison to their monoracial comparison groups. 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level.

coef_table %>%
  filter(mrace=="Black/White" & model=="Raw") %>%
  ggplot(aes(x=term, y=estimate, color=colors))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=lower,
                ymax=upper, fill=colors), color=NA, alpha=0.5)+
  geom_linerange(data=filter(ci_df, mrace=="Black/White" & model=="Raw"), 
                 position=position_dodge2(width = 0.05), size=1.2, alpha=0.5,
                 aes(ymin=lower, ymax=upper, color=colors))+
  geom_point(size=2)+
  coord_flip()+
  scale_y_continuous(labels = scales::percent)+
  scale_color_manual(values=color_choices[c(2,5:6)])+
  scale_fill_manual(values=color_choices[c(2,5:6)])+
  labs(x=NULL,
       y="predicted probability of being behind expected grade")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())
```

Additionally, I include a measure of the "halfway" point between the two racial groups.[^3] This halfway point allows me to determine where the biracial group falls relative to the expectation of being halfway between the two constituent monoracial groups. In this case, Black/White students are much closer to White students in their risk of grade retention, and have probabilities much lower than both Black students and the halfway expectation.

[^3]: The halfway point is given by taking the mean between the two monoracial group. Its standard error is given by $\sqrt{(s_1^2+s_2^2)/2}$ where $s_1$ and $s_2$ are the standard errors of the estimates for the two monoracial groups.

@fig-example shows the raw differences between the three racial groups of interest. However, in practice, I want to estimate differences across groups while holding constant a variety of potentially confounding variables. To do this, I estimate a set of logit models that predict the likelihood of a student being clearly behind expected grade by race and a variety of other variables. I then construct figures similar to @fig-example by calculating from the model the average predicted probabilities (APP) for each racial group. An APP estimates the average probability of the outcome by a given covariate across all cases, while holding constant all other variables. APPs are akin to average marginal effects (AME) and estimated in the same way. Whereas AMEs estimate differences or slopes, APPs estimate the level for a given category.[^4]

[^4]: Output from the full models used to construct these APPs is available in the supplementary materials provided with this article.

I begin with a baseline model that adjusts for a variety of nuisance characteristics that may vary across groups and thus need to be controlled in all models. First, I include fixed effects for state of residence because states dictate educational policy and thus can differ substantially in the likelihood of grade retention. I also include dummy variables indicating whether the student lived in a central city, suburban, or rural area.[^5]

[^5]: The location cannot be determined for all students and I therefore also include a fourth category of unknown in all models.

The probability of being clearly behind expected grade also increases with the student's current grade, so I include fixed effects for the current grade of the student. Additionally, @fig-year-grade shows that the percent of students clearly behind grade level has declined over time, but this decline has been much more substantial at higher grade levels. For the elementary grades, there is no evidence of a decline at all. To account for this grade-specific decline in the models I include an interaction between a linear year term and current grade.[^6]

[^6]: Sensitivity analysis showed that this functional form was preferred by the Bayesian information criterion to models with no interaction and a model with interaction terms that treated year as a categorical variable. I also considered a similar interaction between state and year, but this model was not preferred.

```{r}
#| label: fig-year-grade
#| fig-cap: Trends over time and grade in the percent of students clearly behind grade level. Lines are fit to each set of points by grade via lowess smoothing.

acs %>%
  group_by(current_grade, year) %>%
  summarize(below_grade=mean(below_exp_grade)) %>%
  ggplot(aes(x=year, y=below_grade, group=current_grade, color=current_grade))+
  #geom_line()+
  geom_smooth(se=FALSE)+
  geom_point()+
  scale_x_continuous(breaks=c(2010,2014,2018))+
  scale_y_continuous(labels=scales::percent)+
  labs(x="year", y="percent clearly behind grade level for age",
       color="current grade")+
  theme_bw()
```

The baseline model also includes the race of the respondent. I then add additional terms measuring material and cultural resources that may account for racial differences in a set of subsequent models. First, I include measures of nativity and English proficiency for both the respondent and each of their parents. Second, I include a categorical measure of highest degree earned for each parent. Finally, I include measures of family income (square rooted), home ownership, and whether the parents are married.

All statistics reported here adjust for sampling weights among respondents. All models incorporate design effects for variance in sample weights abd the clustering of multiple respondents within the same household.

# Results

## Monoracial Differences

I begin by showing the relative risk of being behind expected grade among monoracial respondents across models. @tbl-monoracial reports average marginal effects on the probability of being behind expected grade (hereafter BBEG) for each monoracial minority group in comparison to White students. Understanding these differences and how they change across models helps me to understand the potential for how biracial respondents might be positioned.

```{r monoracial-results}
#| results: asis
#| label: tbl-monoracial
#| tbl-cap: Average marginal differences in the probability of being behind grade by monoracial group, relative to a monoracial white student.
texreg(map(results_monoracial[-1], convert_marg),
       digits=3,
       caption = "Average marginal differences in the probability of being clearly behind grade by monoracial group, relative to a monoracial white student",
       caption.above=TRUE,
       label="tbl-monoracial",  #TODO: This does not work
       float.pos="t",
       custom.gof.rows=
         list("State fixed effects" = c("Yes","Yes","Yes","Yes"),
              "Location fixed effects" = c("Yes","Yes","Yes","Yes"),
              "Grade fixed effects" = c("Yes","Yes","Yes","Yes"),
              "Year linear effects" = c("Yes","Yes","Yes","Yes"),
              "Year x grade effects"  = c("Yes","Yes","Yes","Yes"),
              "Nativity and language" = c("No","Yes","Yes","Yes"),
              "Parent's education" = c("No","No","Yes","Yes"),
              "Other family resources" = c("No","No","No","Yes")))
```

Model 1 only controls for structural factors such as state of residence, student location, grade, and year. These results provide a baseline estimate of the differences across monoracial groups without controlling for differences in cultural and material resources across groups. The results show that Black, Indigenous, and Latino students all have substantially higher probability of BBEG than White students. Indigenous students have substantially higher risk than all other students, with a probability of BBEG that is 4.1% percentage points higher than White students. Asian students have the lowest probability of BBEG, and their risk is slightly lower than Whites.

The subsequent models control for a variety of additional variables. Model 2 includes controls for whether the student and each of their parents are native-born and speak English well. Model 3 controls for the highest degree received for each parent, and Model 4 controls for family income (square rooted), home ownership, and whether the student's parents are married.

Model 2 and Model 3 have substantial effects on the observed inequality across racial groups, while the additional variables from Model 4 have less impact. Controlling for the characteristics related to immigration and acculturation in Model 2 cuts the gap between White and Latino students in half and increases the gap between Asian and White students. It has no impact on the gap between White and Black or White and Indigenous students because most of these students are native-born with native-born parents.

Controlling for education in Model 3 reduces the gap substantially for Black and Indigenous students and removes the gap entirely for Latino students. In fact, the results of both Models 3 and 4 suggest that, when holding constant cultural and family resources, Latino students are slightly less likely than White students to be held back.

In total, resource differentials account for a substantial part of the overall racial differences in the risk of BBEG, but not all of these differences. Black and Indigenous students remain at higher risk than other students. Notably, Asian students rather than White students have the lowest risk of BBEG in all models. The most notable shift across models is for Latino students. Their higher risk of BBEG relative to White students is completely accounted for by resource differentials. When comparing hite and Latino students with the same level or resources in Model 4, Latino students actually have slightly lower risk of BBEG.

Although, the racial differences in risk of BBEG are substantially reduced by controlling for resource differentials, the remaining gaps across all pairwise combinations are still large enough in most cases to sustain the question of where biracial individuals fit within these gaps. Before turning to this question, however, I want to better understand how the resources of biracial individuals compare to their monoracial constituent groups.

## The Distribution of Resources for Biracial Individuals

@tbl-resources shows the mean values of the resource variables across monoracial and biracial groups.[^7] Because of the number of variables and groups involved, this table can be overwhelming to read. Therefore, I also provide a summation of these resource differentials by racial group by calculating a counterfactual probability of BBEG for each group based on their observed distribution of resource characteristics. This counterfactual probability is calculated by estimating the average predicted probability for each group from the fullest model (Model 4 of \@tbl-monoracial) when all other variables, including race, are held at their mean.[^8]

[^7]: For compactness, I combine the parent foreign-born and English proficiency questions into whether either parent is foreign-born or speak English. In the models, I use separate variables for mothers and fathers. Tables in the supplementary materials provide the full breakdown of these variables by racial group.

[^8]: It might seem odd to hold categorical variables at their mean, but because all categorical variables are technically entered into the models as 0/1 numeric indicator variables, their mean value can be calculated as the proportion of cases with the indicated value. This technique is frequently used in multilevel models to deal with grand and group mean centering of variables. <!-- TODO CITE raudenbush and bryk? -->

```{r}
#| label: tbl-resources
#| tbl-cap: Mean resources by racial group. All results are survey weighted. Results sorted by counterfactual probability of grade retention. Shading indicates biracial group.
index_biracial <- which(str_detect(tbl_resources$race, "/"))

tbl_resources %>%
  select(race, held_back_cf, fam_inc, own_home, m_ba, f_ba, f_born, par_fb,
         spk_eng_well, spk_eng_par) %>%
  mutate(held_back_cf=percent(held_back_cf/100, accuracy=0.1),
         own_home=percent(own_home/100, accuracy=0.1),
         m_ba=percent(m_ba/100, accuracy=0.1),
         f_ba=percent(f_ba/100, accuracy=0.1),
         f_born=percent(f_born/100, accuracy=0.1),
         par_fb=percent(par_fb/100, accuracy=0.1),
         spk_eng_well=percent(spk_eng_well/100, accuracy=0.1),
         spk_eng_par=percent(spk_eng_par/100, accuracy=0.1),
         fam_inc=dollar(fam_inc)) %>%
  kbl(booktabs = TRUE,
      linesep = NULL, position = "p",
      align = c("l",rep("r", ncol(tbl_resources)-1)),
      col.names=c("Race","Behind expected grade, counterfactual","Family income",
                  "Own home","Mother, 4-year college degree",
                  "Father, 4-year college degree",
                  "Foreign-born","Either parent foreign-born",
                  "Speak English well", 
                  "Both parents speak english well")) %>%
  kable_styling(latex_options = c("scale_down")) %>%
  row_spec(index_biracial, background = "#D3D3D3") %>%
  column_spec(2:10, width = "1.5cm")
#%>%
#  landscape()
```

The results indicate that we would expect very different risk of BBEG across biracial groups because of differences in resources. For example, White/Asian students have the lowest counterfactual risk of BBEG of all groups at 2.4%, owing to their exceptionally high family income and the high college attainment of their parents. At the other end of the spectrum, Indigenous/Latino students have a counterfactual risk of BBEG nearly 2% higher, given their much lower family resources.

My particular interest, however, is in how each biracial group compares to its constituent monoracial groups. For example, while Indigenous/Latino students have the highest counterfactual risk of BBEG among biracial students, their risk is still lower than their two constituent groups of Indigenous (4.5%) and Latino (5.3%) students, owing in large part to a somewhat higher family income, greater parental educational attainment, and greater acculturation. These results suggest positive selection into Indigenous/Latino interracial unions.

```{r}
#| label: tbl-resource-summary
#| tbl-cap: Relative placement for biracial groups in comparison to monoracial constituent groups of counterfactual probability of being behind grade from resources alone.
race <- c("White/Asian","Black/Asian","Latino/Asian","White/Latino",
          "Black/Latino","Indigenous/Latino","White/Indigenous",
          "Black/Indigenous","Indigenous/Latino","Black/White")
lower_both <- c("X",rep("",9))
same_lower <- c("","X (Asian)","X (Asian)","X (White)","X (Black)","X (Indig.)",
                rep("",4))
between <- c(rep("",6),rep("X",3),"")
same_higher <- c(rep("",9),"X (Black)")
higher_both <- rep("",10)
tibble(race, lower_both, same_lower, between, same_higher, higher_both) %>%
 kbl(booktabs = TRUE,
      linesep = NULL,
      align = c("l",rep("r", 5)),
      col.names=c("Biacial","Lower than both","Same as lower","Between",
                  "Same as higher","Higher than both")) %>%
  kable_styling() %>%
  column_spec(1:6, width = "2cm")

```

Table 4 summarizes the patterns from @tbl-resources across all ten biracial groups. In six of these cases, the biracial group would have outcomes similar or better than the monoracial group with better resources. These results imply that for the majority of these groups, parents in these matches are being positively selected from their constituent groups.

Only three groups exhibit an in-between status based on resources. All three of these groups are part-Indigenous. This finding does not hold for Indigenous/Latino students, but nonetheless suggests less selectivity in crossing this boundary.

The Black/White case stands out as an outlier. Black/White students have similar resources as Black students and both groups have fewer resources than White students. This is observable in @tbl-resources across family income, home ownership, and parental educational attainment. Unlike most of these other groups, there does not seem to be substantial positive class selectivity into Black/White interracial unions. 
The effect of controlling for these resources will be complex and different across biracial groups. For those biracial groups with resources similar to their more advantaged constituent group, we should see that these controlling for these resources should reduce their advantage relative to the less advantaged monoracial group. For the Black/White case, on the other hand, we would expect that controlling for resource differentials should eliminate some of their disadvantage relative to Whites. For the remaining “in-between” groups, it is more difficult to make a priori predictions.

## Biracial Placement

I now turn to the results for biracial respondents. Due to the number of comparisons being made, I split these results into two separate figures. In @fig-black-models, I show the placement of each biracial group involving one black parent in comparison to their two monoracial constituent groups. To allow comparison across models, I compare the results from the baseline model (Model 1 in @tbl-monoracial) and the full model that accounts for all cultural and family resources (Model 4 in @tbl-monoracial).

```{r fig-black-models}
#| label: fig-black-models
#| fig-pos: p
#| fig-cap: Probability of being behind grade for biracial respondents with one black parent, in comparison to their monoracial comparison groups. 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level, Baseline model includes year, grade, location, and state fixed effects. Full models include control variables for nativity, English proficiency, income, education, home ownership, and marital status of parents.
#| fig-height: 7.3
coef_table %>%
  filter(model=="Baseline model" | model=="Full model") %>%
  filter(str_detect(mrace, "Black")) %>%
  ggplot(aes(x=term, y=estimate, color=colors))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=lower, ymax=upper, fill=colors), 
            color=NA, alpha=0.5)+
  geom_linerange(data=subset(ci_df, 
                             (model=="Baseline model" |  model=="Full model") &
                               str_detect(mrace, "Black")), 
                 position=position_dodge2(width = 0.1), size=1, alpha=0.5,
                 aes(ymin=lower, ymax=upper, color=colors))+
  geom_point()+
  coord_flip()+
  facet_grid(mrace~model, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=color_choices)+
  scale_fill_manual(values=color_choices)+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average predicted probability of being clearly behind expected grade")
```

I focus first on biracial groups with one Black parent because of a stronger historical expectation that these students will be identified as Black by the "one-drop rule" and as a result will have similar outcomes to Black students. @fig-black-models provides little evidence for such a pattern, and soundly rejects it in the case of Black/White and Black/Latino students. Black/White and Black/Latino students have a risk of BBEG substantially lower than Black students and closer, and perhaps even lower, than their non-Black constituent group. Solid conclusions are difficult for the remaining two groups because of smaller sample size and correspondingly wider confidence intervals on estimates. Nonetheless the point estimates in each of these cases suggest probabilities of BBEG roughly midway between the two constituent groups.

Controlling for resource variables has some effect on the placement of part-Black biracial respondents. In the case of Black/White biracial respondents, Black/White students initially have a risk of BBEG slightly higher than and statistically distinguishable from White students. After controlling for cultural and family resources, however, Black/White respondents have a risk of BBEG that is slightly lower and statistically distinguishable from White students. This change reflects the relatively low level of parental education and income among Black/White students, which is much closer to Black students than White students, as shown on @tbl-descriptive.

An opposite pattern is observed for Black/Latino respondents. Black/Latino respondents initially have a risk of grade retention substantially lower than both Black and Latino students, who have similar risks. However, after controlling for cultural and family resources, Black/Latino students have a risk of BBEG that is indistinguishable from the risk of Latino students and substantially lower than Black students. The change largely reflects the fact that Black/Latino students have greater resources than Latino (but not Black) students and once this advantage is held constant, they have a similar risk.

The role of parental resources is harder to determine for Black/Asian and Black/Indigenous students due to the greater uncertainty in these estimates. Controlling for resources moves the point estimate for Black/Asian students from a point estimate close to Asian students to more of a halfway position between the two constituent groups. Because Black/Indigenous students had resources halfway between their constituent groups, controlling for these resources hardly affects the relative placement of Black/Indigenous students at all.

@fig-other-models shows the results for the remaining biracial groups. In general, the results here show that biracial students are more likely to have a risk of BBEG more like the monoracial constituent group with the lower risk than the higher risk, with some important caveats discussed below.

```{r fig-other-models}
#| label: fig-other-models
#| fig-pos: p
#| fig-cap: Probability of being behind grade for non-black biracial respondents, in comparison to their monoracial comparison groups. 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level, Baseline model includes year, grade, location, and state fixed effects. Full models include control variables for nativity, English proficiency, income, education, home ownership, and marital status of parents.
#| fig-height: 7.3
coef_table %>%
  filter(model=="Baseline model" | model=="Full model") %>%
  filter(!str_detect(mrace, "Black")) %>%
  ggplot(aes(x=term, y=estimate, color=colors))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=lower, ymax=upper, fill=colors), 
            color=NA, alpha=0.5)+
  geom_linerange(data=subset(ci_df, 
                             (model=="Baseline model" | model=="Full model") &
                               !str_detect(mrace, "Black")), 
                 position=position_dodge2(width = 0.1), size=1, alpha=0.5,
                 aes(ymin=lower, ymax=upper, color=colors))+
  geom_point()+
  coord_flip()+
  facet_grid(mrace~model, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=color_choices[-1])+
  scale_fill_manual(values=color_choices[-1])+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average predicted probability of being clearly behind expected grade")
```

For all groups, the full models generally indicate that each biracial group has a risk of BBEG similar to the risk of the lower monoracial constituent group or has risk higher than this group but less than the halfway point. There are two exceptions to this general pattern. First, the point estimate of the probability of BBEG for the Indigenous/Latino group is roughly halfway between the two constituent groups and statistically distinguishable from both. Second, Latino/Asian students are the only case where the point estimate of the risk for the biracial group is higher than either monoracial group. However, the confidence bands suggest that the risk for this group cannot be statistically differentiated from either of the two monoracial constituent groups. This result only emerges in the model that controls for resources, Latino/Asian students come from more advantaged households than both Latino and Asian students, on average.

Aside from the effect on Latino/Asian students, controlling for resources has moderate effects on the placement of the remaining biracial groups. White/Asian students are the only case with a substantial shift, owing to their highly advantaged households. After controlling for resources, the risks of BBEG for White/Asian students are indistinguishable from the risk of Asian students, but still substantially and statistically distinguishable from the higher risk to White students.

Results for part-Latino students are complicated by the substantial change in the monoracial Latino risk across models. For example, while Latinos have substantially higher risk of BBEG than Whites in the baseline model, there is only a very small difference favoring Latinos in the full model. Nonetheless, even in this case, we see that White/Latinos have risk closer to those of the lower Latino group, but distinguishable from both groups.

# Conclusions

-   Big claim: The results are consistent with a new "regime of hyperdescent" operating across all racial groups. This contributes to a muddying of racial distinctiveness via hybridity, while at the same time isolating monoracial members of the most stigmatized racial groups, consistent with arguments like Bonilla-Silva's Latin Americanization.

# Acknowledgments {.appendix}

I am grateful to...

# References {.appendix}