### 2023-03-13 products/paper/supp_materials.qmd
Add table in supplementary materials to address issue #33

---
title: "Supplementary Materials for Differences in the Risk of Grade Retention for Biracial and Monoracial Students in the US, 2010-2019"
format:
  pdf:
    documentclass: scrartcl
    include-in-header: 
      text: |
        \usepackage{longtable}
    papersize: letter
    fig-height: 4
    fig-pos: h
    keep-tex: false
  docx:
    toc: false
bibliography: ../project.bib
execute: 
  eval: true
  echo: false
  warning: false
  error: false
prefer-html: true
editor: source
csl: american-sociological-association.csl
---

```{r}
#| label: setup
#| include: false
library(here)
source(here("analysis","check_packages.R"))
source(here("analysis","useful_functions.R"))
load(here("analysis","output","bic_spec_tests.RData"))
load(here("analysis","output","model_summaries.RData"))
load(here("analysis","output","marg_prob_alt.RData"))
load(here("analysis","output","acs.RData"))
```

```{r}
#| label: func-convert-model
#| include: false

#  because I only saved model summaries I need a function to convert this to 
# something that knitreg likes
convert_model <- function(model) {
  tr <- createTexreg(
    coef.names = rownames(model$coefficients),
    coef = model$coefficients[,"Estimate"], 
    se = model$coefficients[,"Std. Error"], 
    pvalues = model$coefficients[,"Pr(>|t|)"],
    gof.names = c("N","Null deviance","Model deviance", "Pseudo-R2"), 
    gof = c(model$df.null+1, model$null.deviance, model$deviance, 
            (model$null.deviance-model$deviance)/model$null.deviance), 
    gof.decimal = c(F,F,F,T)
  )
}
```

```{r}
#| label: create-design
#| eval: false
acs_design <- acs %>%
  as_survey_design(ids = cluster, weights = perwt)
```

```{r}
#| results: asis
#| label: tbl-full-models
#| tbl-cap: Log-odds ratios from full models predicting a student being clearly behind expected grade. All models used sample weights and adjust for design effects from sample weight variance and clustering of students within the same household.
texreg(lapply(list(model0_summary, model1_summary, model2_summary, 
                   model3_summary, model4_summary), convert_model),
       digits = 3,
       longtable = TRUE,
       use.packages = FALSE,
       caption.above = TRUE,
       #file="full_reg.doc",
       custom.coef.names = 
         c("Intercept","Black","Indigenous","Asian","Latino","Black/White",
           "Black/Indigenous","Black/Latino", "Black/Asian",
           "White/Indigenous","White/Latino","White/Asian",
           "Indigenous/Latino","Indigenous/Asian","Latino/Asian",
           "Year (2010 origin)", "1st grade","2nd grade","3rd grade",
           "4th grade","5th grade","6th grade","7th grade","8th grade",
           "9th grade","10th grade","11th grade","12th grade","Alaska",
           "Arizona","Arkansas","California","Colorado","Connecticut",
           "Washington DC","Delaware","Florida","Georgia","Hawaii","Idaho",
           "Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana",
           "Maine","Maryland","Massachusetts","Michigan","Minnesota",
           "Mississippi","Missouri","Montana","Nebraska","Nevada",
           "New Hampshire","New Jersey","New Mexico","New York",
           "North Carolina","North Dakota","Ohio","Oklahoma","Oregon",
           "Pennsylvania","Rhode Island","South Carolina","South Dakota",
           "Tennessee","Texas","Utah","Vermont","Virginia","Washington",
           "West Virginia","Wisconsin","Wyoming","Non-metro","Central city",
           "Metro non-central","Metro indeterminable","Female",
           "Year x 1st grade","Year x 2nd grade","Year x 3rd grade",
           "Year x 4th grade","Year x 5th grade","Year x 6th grade",
           "Year x 7th grade","Year x 8th grade","Year x 9th grade",
           "Year x 10th grade","Year x 11th grade","Year x 12th grade",
           "Foreign-born","Mother foreign-born","Father foreign-born",
           "Speaks English somewhat", "Speaks English well", 
           "Mother speaks somewhat", "Mother speaks well",
           "Father speaks somewhat", "Father speaks well",
           "Mother High school diploma", "Mother AA degree", 
           "Mother BA degree","Mother Grad degree",
           "Father High school diploma", "Father AA degree", 
           "Father BA degree","Father Grad degree",
           "Family income (10K USD, sq. root)", "Owns home",
           "parents married"),
       groups = list("Race (ref. White)"=2:15, 
                     "Grade (ref. Kindergarten)"=17:28,
                     "State (ref. Alabama)"=29:78,
                     "Location (ref. Indeterminable)"=79:82,
                     "English proficiency (ref. none)"=99:100,
                     "Mother's English proficiency (ref. none)"=101:102,
                     "Father's English proficiency (ref. none)"=103:104,
                     "Mother's highest degree (ref. none)"=105:108,
                     "Father's highest degree (ref. none)"=109:112))
```

```{r tbl-bic}
#| label: tbl-bic
#| tbl-cap: Bayesian Information Criterion (BIC) statistics for models accounting for change over time using different specifications. All models also include categorical dummy variables for race, grade, state, and location.
bic_scores$model[1] <- "Categorical year dummies, no interaction"
bic_scores$model[2] <- "Categorical year dummies, year * grade interaction"
bic_scores$model[3] <- "Continous year variable, no interaction"
bic_scores$model[4] <- "Continous year variable, year * grade interaction"
bic_scores$model[5] <- "Continous year variable, year * grade interaction + year * state interaction"

bic_scores <- bic_scores[c(1,3,2,4,5),]

bic_scores %>%
  kbl(booktabs = TRUE,
      linesep = NULL,
      align = c("l","r"),
      col.names=c("Model", "BIC Score")) %>%
  kable_styling()
```

```{r marg-fit-table}
#| label: tbl-marg-fit
#| tbl-cap: Marginal probabilities of grade retention by racial group across models using different specifications to model the the temporal trend in grade retention by current grade of the student.
marg_prob_alt %>%
  mutate(baseline=percent(round(baseline, 5)),
         year_dummy=percent(round(year_dummy, 5)),
         year_linear=percent(round(year_linear, 5)),
         year_dummy_interact=percent(round(year_dummy_interact, 5)),
         year_linear_interact=percent(round(year_linear_interact, 5)),
         year_grade_state=percent(round(year_grade_state, 5))) %>%
  kbl(booktabs=TRUE, linesep=NULL,
      align = c("l",rep("r", ncol(marg_prob_alt)-1)),
      col.names=c("Race", "No Covariates", "Year as dummy", 
                  "Year as linear", "Year (dummy) x Grade",
                  "Year (linear) x Grade", 
                  "Year (linear) x Grade and State")) %>%
  add_header_above(c(" " = 2, 
                     "Sensitivity tests on calendar year x grade coding" = 5)) %>%
  column_spec(2:7, width = "1.75cm") %>%
  footnote(general = "All models except no covariate case include dummies for current grade and state")
```

```{r desc-stats-ses}
#| label: tbl-desc-stats-ses
#| tbl-cap: Socioeconomic resources by race. Shading indicates biracial group. All results apply survey weights.
#| eval: false
tbl_summary <- acs_design %>%
  group_by(race) %>%
  summarize(fam_inc=survey_mean(family_income, vartype = NULL))

tbl_summary <- acs_design %>%
  group_by(race) %>%
  summarize(fam_inc=round(survey_mean(family_income, vartype = NULL), 0),
            own_home=round(survey_mean(own_home, vartype = NULL), 3)*100,
            par_mar=round(survey_mean(parents_married, vartype = NULL), 3)*100,
            f_lhs=round(survey_mean(degree_father=="LHS", vartype = NULL), 
                        3)*100,
            f_hs=round(survey_mean(degree_father=="HS", vartype = NULL), 3)*100,
            f_aa=round(survey_mean(degree_father=="AA", vartype = NULL), 3)*100,
            f_ba=round(survey_mean(degree_father=="BA" | degree_father=="G", 
                                   vartype = NULL), 3)*100,
            m_lhs=round(survey_mean(degree_mother=="LHS", vartype = NULL), 3)*100,
            m_hs=round(survey_mean(degree_mother=="HS", vartype = NULL), 3)*100,
            m_aa=round(survey_mean(degree_mother=="AA", vartype = NULL), 3)*100,
            m_ba=round(survey_mean(degree_mother=="BA" | degree_mother=="G", 
                                   vartype = NULL), 3)*100)

index_biracial <- which(str_detect(tbl_summary$race, "/"))

tbl_summary %>%
  mutate_if(is.numeric, format, big.mark=",") %>%
  kbl(booktabs=TRUE, linesep=NULL,
      align = c("l",rep("r", ncol(tbl_summary)-1)),
      col.names=c("race","mean family income","% own home",
                  "% parents married",
                  "% father, no HS diploma","% father, HS diploma",
                  "% father, AA degree",  "% father, BA+ degree",
                  "% mother, no HS diploma","% mother, HS diploma", 
                  "%  mother, AA degree", "% mother, BA+ degree")) %>%
  kable_styling(font_size=8) %>%
  column_spec(2:12, width = "0.75cm") %>%
  row_spec(index_biracial, background = "#D3D3D3")
```

```{r}
#| label: tbl-desc-stats-cult
#| tbl-cap: Cultural resources by race. Shading indicates biracial group. All results apply survey weights.
#| eval: false
tbl_summary <- acs_design %>%
  group_by(race) %>%
  summarize(f_born=round(survey_mean(foreign_born, vartype = NULL), 3)*100,
            father_fb=round(survey_mean(foreign_born_father, 
                                        vartype = NULL), 3)*100,
            mother_fb=round(survey_mean(foreign_born_mother, 
                                        vartype = NULL), 3)*100,
            spk_eng_no=round(survey_mean(spk_eng=="No", vartype = NULL), 3)*100,
            spk_eng_some=round(survey_mean(spk_eng=="Somewhat", 
                                           vartype = NULL), 3)*100,
            spk_eng_yes=round(survey_mean(spk_eng=="Yes", vartype = NULL), 3)*100,
            f_spk_eng_no=round(survey_mean(spk_eng_father=="No", 
                                           vartype = NULL), 3)*100,
            f_spk_eng_some=round(survey_mean(spk_eng_father=="Somewhat", 
                                             vartype = NULL), 3)*100,
            f_spk_eng_yes=round(survey_mean(spk_eng_father=="Yes", 
                                            vartype = NULL), 3)*100,
            m_spk_eng_no=round(survey_mean(spk_eng_mother=="No", 
                                           vartype = NULL), 3)*100,
            m_spk_eng_some=round(survey_mean(spk_eng_mother=="Somewhat", 
                                             vartype = NULL), 3)*100,
            m_spk_eng_yes=round(survey_mean(spk_eng_mother=="Yes", 
                                            vartype = NULL), 3)*100)

index_biracial <- which(str_detect(tbl_summary$race, "/"))

tbl_summary %>%
  mutate_if(is.numeric, format, big.mark=",") %>%
  kbl(booktabs=TRUE, linesep=NULL,
      align = c("l",rep("r", ncol(tbl_summary)-1)),
      col.names=c("race",
                  "% foreign-born","% father foreign-born",
                  "% mother foreign-born",
                  "% speak no English", "% speak some English", 
                  "% speak English",
                  "% father speak no English", "% father speak some English", 
                  "% father speak English",
                  "% mother speak no English", "% mother speak some English",
                  "% mother speak English")) %>%
  kable_styling(font_size=8) %>%
  column_spec(2:13, width = "0.75cm") %>%
  row_spec(index_biracial, background = "#D3D3D3")
```