### 2023-03-14 products/submissions/SocSci/randr/response_memo.md
Add submission materials for SocSci R&R

# Response Memo, Sociological Science

Thank you to the reviewer and editor for the helpful comments and to the editor for the opportunity to revise my manuscript. I have made changes to the manuscript as requested and I believe those changes have improved the overall manuscript. Those changes are described below.

## Issue 1

From the editor's letter:

> Please provide descriptives on the distribution of CBEG across racial groups and discuss the (im-)possibility of false positives and how this could (not) have impacted the results. DE: “What is the distribution of this measure across across groups? Some descriptives here would help us. For example, do some groups systematically enroll their kids in school later than others? In other words, is the measure truly capturing what we want it to capture? (For example, in Figure 2, we see that it is possible to have a CBEG status even in kindergarten. What is keeping these students behind? Or are they simply starting later than others?)”

There are a couple of points to address here. First, I believe the concern that the reviewer is raising, although they do not name it directly, is the phenomenon of "academic redshirting" in which parents hold back kindergarten enrollment for a child whose birthday falls close to the cutoff date for enrollment. It is true that I should have addressed this issue more directly in the paper and I have done so in the revised draft.

* The distribution of CBEG across racial groups was already provided in the first draft in table 1. However, I think what the reviewer is interested in is the distribution of CBEG across racial groups and grades with a particular focus on kindergarten. I have added a table to the supplementary appendix (Table 2) that shows the probability of CBEG jointly across race and current grade. There is some evidence at the kindergarten level of the usual racial disparities being reversed. Most notably, Black and Latino kindergarteners are less likely to be CBEG than White kindergartners. This finding is consistent with prior work on the topic of redshirting (Bassok and Reardon 2013). However, these inverted racial differences quickly flip at higher grades.
* Based on the belief that redshirting is more likely to affect results at younger grades where it is a larger share, I ran a sensitivity analysis that repeated my full models but separately for the elementary grades w/o kindergarten (1st-5th), middle school (6th-8th), and high school (9th-12th). The results from this sensitivity analysis are provided in Figure 1 of the supplementary materials. Although somewhat noisier, they are consistent with the patterns found in the main text. This sensitivity analysis is described in footnote 1 of the main text which reads:

> Analysis of the probability of being CBEG by race and grade, shown in the supplementary materials, indicates that Black and Latino children in kindergarten are less likely to be CBEG than White children, which may be a consequence of redshirting. However, these differences quickly reverse direction by first grade. As a sensitivity analysis, I repeated the main analysis shown here separately for students in elementary (1st-5th), middle school (6th-8th), and high school (9th-12th) grades. These results are presented in the supplementary materials. Although more statistically noisy due to smaller sample sizes, those results are consistent with the conclusions drawn here.

* I added a discussion of redshirting to the main body of the text on pg. 13-14 of the main text which reads:

> Using grade-age comparisons to infer grade retention may inadvertently capture cases of “academic redshirting” in which parents intentionally delay their child’s kindergarten enrollment by a year. Frederick and Hauser (2008) suggest that academic redshirting is in many cases be a form of preemptive retention for children with developmental delays and thus the error induced by these false positive cases may be minimal. However, Bassok and Reardon (2013) have shown that academic redshirting is more common among white and SES privileged parents, suggesting that the demographic covariates of academic redshirting may operate in the opposite direction of grade retention. Regardless, academic redshirting should be of minimal concern for the measure of CBEG used here. Academic redshirting most frequently occurs for students whose birthdays fall close to the cutoff period for enrollment (Bassok and Reardon 2013; Graue and DiPerna 2000). Therefore, academic redshirts spend most of a given school year at the older, but correct, modal age for their grade and most of these students will not be identified as CBEG.

## Issue 2

From the editor's letter:

> Please clarify whether results remain unaltered when year dummies are used. DE: “One question about the specification: The author notes that linear year term interacted with current grade was preferred by the Bayesian Information Criterion in model comparison. But did the main results remain similar across different specifications (e.g., year dummies, year dummies x current grade)? It would be good to include more robustness checks to ensure the results are not driven by any of the decisions made along the way.”

This is a good point. I have now added a new table (Table 4) to the supplementary materials which shows the marginal probabilities of CBEG for each racial group across all the different specifications of year, current grade, and state that were considered. These probabilities are virtually identical down to the fifth decimal place across specifications. I discuss this analysis in footnote 5 of the main text.

## Issue 3

From the editor's letter:

> The paper is now framed as providing the first good evidence on the matter at hand, saying previous studies had two forms of data issues. But previous studies did have some significant findings despite smaller samples, and these findings seem to deviate from the ones obtained here, specifically Herman (2009) and Campbell (2009). Consider whether the paper’s impact would increase if you engaged with these prior findings more, perhaps presenting the results as in contrast with earlier findings, and considering how they may be reconciled. The paper could also engage more with findings from earlier studies using identity measures of ethic/racial group membership, and more clearly articulate how the issue of reverse causality solved in the present paper reconciles differences in findings. This may help make clear how the paper moves the field forward. This could all be done in the discussion section, which currently present the results on their own, without relating them to earlier findings.

Thank you for the nudge on this. My final three paragraphs of the conclusions were meant to address but they did so indirectly. I have now edited those paragraphs to discuss the prior work and how that relates to both the strengths and weaknesses of the current work and pathways for future work. It now reads as follows:

> These findings stand in contrast to earlier work which found more evidence of an “in-between” status in other educational outcomes (Campbell 2009; Herman 2009). I raise two possible explanations for this discrepancy that point to strengths and weaknesses of the current research and may help drive future work on the topic. First, this discrepancy may be due to the small sample sizes of multiracial respondents in prior work which led to statistically imprecise estimates. Because the current findings rely upon large samples, they present an important step forward in our understanding of how biracial individuals will fit into America’s system of racial inequality. Nonetheless, sample sizes for some non-white biracial respondents remain small, limiting our understanding of the outcomes for those groups. In particular, results for non-white part-Indigenous populations are quite imprecise. The results counter-intuitively suggest that part-Indigenous populations are the least likely to be in the position expected by hyperdescent, but due to low sample sizes, that finding is highly tentative.

> Second, The current findings may also diverge from the findings of prior work because different outcomes were examined. While grade retention is an important early life outcome, it is far from the only one that these individuals will face in their life. Although, we often observe similar racial inequalities across a variety of outcomes, knowing the result for one outcome does not perfectly predict other outcomes. As biracial individuals age and face later outcomes in the educational system, the labor market, and elsewhere, their experiences may differ from what I observe here. Thus, collecting more detailed information on these outcomes in future studies remains critically important.